---
title: "Batteries"
subtitle: "Car Fires"
output: html_document
date: "2025-07-25"
---

# Cars Fires

## Action Taken

```{r setup-01a, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)

library(magrittr)
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(rstan)
# library(RPostgreSQL)

keep <- c(ls(), "figs")

if(! exists("csv_dir")){
  csv_dir <- "csv/"
}
if(! exists("intmdt_dir")){
  intmdt_dir <- "rdata/"
}

source("load.files.R")
load.data(final=c("CarFires.RData", "stan.models.RData"), 
          data =c("CarFires.data.RData", "stan.models.RData"), 
          dir=intmdt_dir)
rm(load.data)

keep2 <- c("car.fires", "tesla", "test.cases", "results", "n.est", "car.actions",
           "nfpa", "stan.tesla", "stan.slope", "figs")
```

This looks at action taken for Li ion vehicle fires to see if there is anything 
useful there.

```{r tesla-action, eval=FALSE}
action.codes <- read.csv(paste0(csv_dir, "act_tak_codes.csv"))[,-1]
car.actions[,11:13] %>% as.matrix() %>% as.vector() %>% {.[! is.na(.) & . != '' & . != 'NULL']} %>%
           data.frame(code=.) %>% left_join(action.codes, by=join_by(code==code_value)) %>%
           with(table(code_descr)) %>% sort(decreasing=T)
```

Action taken does not give anything useful. Basically, the only thing it says is
"Put out fire" as the action taken. Not useful.

## Fire Estimates

```{r start-data}
################################################################################
########### This Section Prepares Data and Models                    ###########
################################################################################
```


```{r tesla-fires, eval=! exists("tesla")}
#The data used here are found at this site:
# "https://www.tesla-fire.com/"
# 
# The data was downloaded and imported for use.

tesla <- paste0(csv_dir, "Tesla Fire.xlsx") %>% openxlsx2::wb_to_df(1) %>% 
         { setNames(., c("ID", "Year", "Date", "Country", "State", "Model", "Fatalities", "Description", 
                         paste0("source", 1:(ncol(.) - 8)))) } %>% 
         filter(! is.na(Year)) %>% 
         select(ID, Year, Date, Country, State, Model, Fatalities, Description) %>%
         mutate(Date=lubridate::ymd(Date))
```


```{r nfpa-fires, eval=! exists("nfpa")}
#The data used here is downloaded from the NFPA
# 
# The data was downloaded and imported for use.
nfpa <- openxlsx2::read_xlsx(paste0(csv_dir, "new_sales.xlsx"), 4)
nfpa <- car.summ %>% full_join(nfpa %>% select(Year, Fires, Deaths, Injuries),
                       by=join_by(year==Year)) %>% 
             setNames(c("Year", "NFIRS", "VINs", "NFPA", "Deaths", "Injuries"))
rm(car.summ)
```

```{r vin-lookup, eval=! exists("test.cases")}
car.fires$inc_date <- lubridate::ymd(car.fires$inc_date)
car.fires$inc_type <- as.character(car.fires$inc_type)

car.fires$mob_vin_no[is.na(car.fires$mob_vin_no)] <- "NA"
car.fires$mob_vin_no <- car.fires$mob_vin_no %>% 
                                toupper() %>% 
                                str_replace_all("I", "1") %>% 
                                str_replace_all("O", "0") %>% 
                                str_pad(17, "right", "*")

reg.check.vin <- data.frame(vin=car.fires$mob_vin_no) %>% 
                      mutate(r0=str_detect(vin, "[ -]|^[0-9*]+$", negate=T),
                             r1=str_detect(vin, "^[0-9A-HJ-NPR-Z*]{12}"),
                             r2=str_detect(vin, "[0-9*]{5}$"),
                             r3=str_detect(vin, "^[0-9A-Z*]{8}([0-9X])"),
                             r4=str_detect(vin, "^[0-9A-Z*]{9}([^0Z])"))
# Check digit in VIN
coded <- data.frame(text=c(0:9, LETTERS, "*"),
                    value=c(0:9, 1:9, 1:9, 2:9, NA))
coded$value[coded$text %in% c('I', 'O' , 'Q')] <- NA
multiplies <- c(8:2, 10, 0, 9:2)

tmp <- str_split_fixed(reg.check.vin$vin, "", 17)
tmp <- coded$value[match(tmp, coded$text)] %>% matrix(ncol=17) %>% multiply_by_matrix(multiplies) %>%
       magrittr::mod(11)

reg.check.vin$r5 <- substring(reg.check.vin$vin, 9, 9) == c(0:9, 'X', '@')[ifelse(is.na(as.integer(tmp)), 11, as.integer(tmp)) + 1]

ndx <- with(reg.check.vin, r0 & r1 & r2 & r3 & r4)
test.cases <- data.frame(vin=str_pad(toupper(car.fires$mob_vin_no), 17, "right", "*")[ndx],
                         year=car.fires$mob_year[ndx],
                         CheckValid=reg.check.vin$r5[ndx]) %>% 
                  mutate(wmi=substring(vin, 1, 3),
                         wmi=ifelse(substring(wmi, 3, 3) == '9', paste0(wmi, substring(vin, 12, 14)), wmi),
                         rest=substring(vin, 4,17)) %>% unique()

dbExecute(con, "truncate \"tmpVinCases\"")
dbWriteTable(con, "tmpVinCases", test.cases, row.names=F, append=T)

out.cases <- dbGetQuery(con, "SELECT * FROM \"GetElements\"")

test.cases <- out.cases %>% 
       group_by(vin, ModelYear, InputYear, CheckValid, VehicleType, nMakes, Make0, Element) %>% 
       summarize(value=first(Value)) %>%
       pivot_wider(id_cols=1:7, names_from=Element, values_from=value) %>% 
       mutate(`NA`=NULL, 
              Engine=`Engine Model`, 
              Fuel1=`Fuel Type - Primary`,
              Fuel2=`Fuel Type - Secondary`,
              Electric=`Electrification Level`,
              Battery=`Battery Type`,
              .keep="unused")
```

```{r car-fire-types, eval=! "Tesla" %in% names(car.fires)}
car.fires$Tesla <- str_detect(car.fires$mob_vin_no, '^(5YJ|7SA)')
car.fires$BEV   <- car.fires$mob_vin_no %in% test.cases$vin[str_detect(test.cases$Electric, 'BEV')]
car.fires$Plugin<- car.fires$mob_vin_no %in% test.cases$vin[str_detect(test.cases$Electric, 'BEV|PHEV')]
car.fires$Hybrid<- car.fires$mob_vin_no %in% test.cases$vin[str_detect(test.cases$Electric, 'BEV|HEV')]
```


```{stan v2, output.var="stan.model", eval=FALSE}
data {
  int<lower=0>       S1;         // Number of cars in sample 1
  int<lower=0>       S2;         // Number of cars in sample 2
  int<lower=0>       S12;        // Number of cars in both samples
  real               alpha;
  real               scale_p;
}

transformed data {
  int      which;
  real     theta;
  real     min_N;
  real     scale_N;

  if(S1 <= S2){
    which = 1;
    theta = alpha ^ (1. / S2);
  } else {
    which = 2;
    theta = alpha ^ (1. / S1);
  }
  if(S12 == 0){
    scale_N = sqrt(S1 * S2 / ((1 - alpha ^ (1. / S1)) * (1 - alpha ^ (1. / S2))));
  } else {
    scale_N = S1 * S2 / S12;
  }
  min_N = max(S1, S2) / scale_N;
}

parameters {
  vector<lower=0, upper=1>[2]   p;
  real<lower=min_N>             N_hat;
}

transformed parameters {
  real    N = N_hat * scale_N;
}

model {

  target += lgamma(N + 1) - lgamma(S1 + 1) - lgamma(N - S1 + 1)
            + S1 * log(p[1]) + (N - S1) * log(1 - p[1]);
  target += lgamma(N + 1) - lgamma(S2 + 1) - lgamma(N - S2 + 1)
            + S2 * log(p[2]) + (N - S2) * log(1 - p[2]);
  if(S12 == 0){
    theta ~ beta((1 - p[which]) * scale_p, p[which] * scale_p);
  } else {
    target += lgamma(S1 + 1) - lgamma(S12 + 1) - lgamma(S1 - S12 + 1)
              + S12 * log(p[2]) + (S1 - S12) * log(1 - p[2]);
  }
}

generated quantities {
}
```

```{stan stan-slope-v, output.var="stan.mode1", eval=FALSE}
// This differs from the version above in that it computes a slope over time
// instead of computing estimated values for each year separately.
data {
  int<lower=1>          T;       // Number of time periods
  vector[T]             S1;      // Sample 1 Counts
  vector[T]             S2;      // Sample 2 Counts
  vector[T]             So;      // Sample Overlap Counts
  int<lower=0, upper=1> slope1;  // 0: No; 1: Yes
  int<lower=0, upper=1> slope2;  // 0: No; 1: Yes
}

transformed data {

  matrix[T,3]    dats = append_col(append_col(S1, S2), So);
  
  real     min_N;
  real     scale_N = 0;
  int      k=0;
  for(i in 1:T){
    if(So[i] > 0){
      scale_N += S1[i] * S2[i] / So[i];
      k = k + 1;
    }
  }
  scale_N = scale_N / k;
  
  min_N = max(append_row(S1, S2)) / scale_N;
}

parameters {
// The affine transformations make sure that Stan's initial sampling occurs
// near the actual expected values. Otherwise bad initial values become too
// common to reliably come up with results.
  vector<offset=-2>[1 + slope1]  p1_lgt;
  vector<offset=-2>[1 + slope2]  p2_lgt;
  real<multiplier=0.25>          r;
  real<lower=min_N>              N_hat;
}

transformed parameters {
  vector[T]    N;
  vector[T]    p1;
  vector[T]    p2;
  for(i in 1:T){
    N[i] = N_hat * scale_N * exp((i - 1) * r);
    if(slope1 == 1){
      p1[i] = inv_logit(p1_lgt[1] + (i - 1) * p1_lgt[2]);
    } else {
      p1[i] = inv_logit(p1_lgt[1]);
    }
    if(slope2 == 1){
      p2[i] = inv_logit(p2_lgt[1] + (i - 1) * p2_lgt[2]);
    } else {
      p2[i] = inv_logit(p2_lgt[1]);
    }
  }
}

model {
  matrix[T, 3]   p_hat = append_col(append_col(p1, p2), p1 .* p2);
  for(i in 1:T){
    for(j in 1:3){
      target += lgamma(N[i] + 1) - lgamma(dats[i,j] + 1) - lgamma(N[i] - dats[i,j] + 1)
              + dats[i,j] * log(p_hat[i,j]) + (N[i] - dats[i,j]) * log(1 - p_hat[i,j]);
  }}
}

generated quantities {
}
```

```{r start-results}
################################################################################
########### Results Start Here                                       ###########
################################################################################
```


```{r car-fire-est, eval=! exists("results")}
# This section just does the basic Capture-Recapture calculation
# (after building the data tables, of course).
results <- list()
for(i in c(0, 4, 7)){
  nm <- paste0("Both", i)
  results[[nm]] <- tesla %>% filter(tolower(Country) == 'usa') %>% 
                             inner_join(car.fires %>% filter(Tesla) %>% 
                                          mutate(ldate=inc_date - i, hdate=inc_date + i),
                                        by=join_by(Date >= ldate, Date <= hdate, State==state)) %>% 
                             mutate(valid=T) %>% 
                             select(ID, Date, inc_date, State, city, fd_name, Model, mob_model, 
                                    mob_year, ignited, burned, Fatalities, deaths, 
                                    mob_vin_no, inc_type, prop_use, valid, Description) %>% 
                             unique() %>% arrange(ID)
}

######################################################################
# The `Both7` cases are reviewed by hand to determine whether they
# are valid matches. The vector below is the product of that matching.

valid <- c(F, T, T, T, T, F, F, T, 
           F, T, F, F, F, F, F, F, 
           F, F, T, T, F, F, F, T, 
           T, T, F, T, F, T, T)

results$Both7$valid <- valid
rm(valid)
######################################################################

results$Summary <- 
    full_join(tesla %>% filter(tolower(Country) == "usa") %>% group_by(Year) %>% 
                        summarize(Total=n(), FL=sum(State == 'FL'), `No FL`=sum(State != 'FL')) %>% 
                        pivot_longer(2:4, names_to="State", values_to="Tesla"),
              rbind(car.fires %>% mutate(Year = lubridate::year(inc_date), State="Total") %>% 
                             filter(Year > 2005) %>% group_by(State, Year) %>% 
                             summarize(t.fires=sum(Tesla), BEV=sum(BEV), 
                                       Plugin=sum(Plugin), Hybrid=sum(Hybrid)),
                    car.fires %>% mutate(Year = lubridate::year(inc_date), 
                                         State=ifelse(state == 'FL', 'FL', 'No FL')) %>% 
                             filter(Year > 2005, ! is.na(State)) %>% group_by(State, Year) %>% 
                             summarize(t.fires=sum(Tesla), BEV=sum(BEV), 
                                       Plugin=sum(Plugin), Hybrid=sum(Hybrid))))

for(i in c("Both0", "Both4", "Both7")){
  results$Summary <- left_join(results$Summary,
                               results[[i]] %>% mutate(Year=lubridate::year(Date)) %>% 
                                                filter(valid) %>% group_by(Year) %>% 
                                                summarize(Total=n(), 
                                                          FL=sum(State == "FL"), 
                                                          `No FL`=sum(State != 'FL')) %>% 
                                                pivot_longer(2:4, names_to="State", values_to=i))
  results$Summary[[i]][is.na(results$Summary[[i]])] <- 0
}

results$Summary <- rbind(results$Summary,
                         results$Summary %>% filter(Year < 2024) %>% group_by(State) %>%
                             summarize(across(.cols=c("Tesla", "t.fires", "BEV", "Plugin", 
                                                      "Hybrid", "Both0", "Both4", "Both7"), 
                                              .fns=function(x) sum(x, na.rm=T))) %>% 
                             cbind(Year=NA, .))

results$Summary <- results$Summary %>% 
              mutate(#tesla0  = Tesla * t.fires / Both0,
#                     tesla4  = Tesla * t.fires / Both4,
                      tesla7  = Tesla * t.fires / Both7,
#                     bev0    = Tesla * BEV / Both0,
#                     bev4    = Tesla * BEV / Both4,
                      bev7    = Tesla * BEV / Both7,
#                     plugin0 = Tesla * Plugin / Both0,
#                     plugin4 = Tesla * Plugin / Both4,
                      plugin7 = Tesla * Plugin / Both7
                     ) %>% arrange(State, Year)
```

Looking at the data, Florida is dramatically different than the rest of the 
states, with a very high degree of overlap between the `TeslaFires.com` dataset
and NFIRS. Since Florida is very different from the rest of the country in terms
of the overlap between the data sets, the analysis was run twice. First, the 
entire dataset was analyzed as a whole. Second, Florida and the rest of the 
country were analyzed separately.

```{r car-fire-table}
results$Summary[,-c(7:9)] %>% knitr::kable(digits=0)
```

The table above lists capture-recapture estimates for the number of Tesla, EV, 
and Plugin-Hybrid vehicle fires. Note that the last column (the 'Plugin7' column)
represents the total of EV and plugin-hybrid fires. The final line (with no 
year) represents the estimated total for all years in the sample. In addition,
Total results are reported as well as the split results for Florida and the 
rest of the country.

```{r run_stan1, eval=! exists("stan.tesla")}
n.est <- results$Summary %>% filter(is.na(Year) | Year < 2024) %>% 
         select(Year, State, Tesla, t.fires, BEV, Plugin, Both7) %>% 
         pivot_longer(c(t.fires, BEV, Plugin)) %>% 
         rename(Both=Both7) %>% 
         mutate(Tesla=as.integer(Tesla),
                Both=as.integer(Both),
                name=ifelse(name == 't.fires', "Tesla", name),
                value=as.integer(value),
                est=as.numeric(NA),
                lcl=est,
                ucl=est,
                med=est)
stan.tesla <- list()
nm <- paste0(n.est$State, ":", n.est$name, ":", ifelse(is.na(n.est$Year), "Total", n.est$Year))
vv <- with(n.est, Tesla * Both * value)
for(i in 1:nrow(n.est)){
  if(with(n.est, is.na(Year[i]) | Year[i] > 2018) & vv[i] > 0){
    st.data <- with(n.est, list(S1=Tesla[i], S2=value[i], 
                                S12=ifelse(is.na(Both[i]), 0, Both[i]),
                                alpha=0.1, scale_p=10))
    stan.tesla[[nm[i]]] <- sampling(stan.model, data=st.data, refresh=0)
    n.est[i,7:10] <- stan.tesla[[nm[i]]] %>% extract("N") %>% magrittr::extract2(1) %>% 
                    {c(mean(.), quantile(., c(0.025, 0.975, 0.5)))} %>% t()
  }
}
rm(i, nm, vv)
```

```{r run_stan2, eval=! exists("stan.slope")}
stan.slope=list()
ndx <- results$Summary$Year %in% 2019:2023
for(i in c("Total", "FL", "No FL")){
  ndy <- results$Summary$State == i
  for(j in c("t.fires", "BEV", "Plugin")){
    st.data <- list(T=2023 - 2018, 
                    S1=results$Summary$Tesla[ndx & ndy], 
                    S2=results$Summary[[j]][ndx & ndy], 
                    So=ifelse(is.na(results$Summary$Both7[ndx & ndy]), 0, 
                              results$Summary$Both7[ndx & ndy]),
                    slope1=0, 
                    slope2=0)
    nm <- paste0(i, ":", sub("t.fires", "Tesla", j))
    stan.slope[[nm]] <- sampling(stan.mode1, data=st.data, refresh=0)
  }
}
```


```{r more-n-est}
# Add the slope into the n.est data frame
tmp <- stan.slope %>% lapply(function(y) y %>% extract("N") %>% magrittr::extract2(1) %>% 
                                  apply(2, function(x) c(mean(x), quantile(x, c(0.025, 0.975, 0.5)))) %>% 
                                  t() %>% as.data.frame() %>% 
                                 setNames(c("slope", "s.lcl", "s.ucl", "s.med"))) 

for(i in names(tmp)){
  nm <- strsplit(i, ":")[[1]]
  tmp[[i]] <- cbind(Year=2019:2023,
                    State=nm[1],
                    name=nm[2],
                    tmp[[i]])
}

n.est <- left_join(n.est, do.call("rbind", tmp))
rm(tmp, i, nm)
```

```{r fires-fig}
if(! exists("figs")) figs <- list()
figs$ev <- n.est %>% filter(! is.na(est) & ! is.na(Year) & name != "Tesla") %>% 
                ggplot() + aes(x=Year, y=est, ymin=lcl, ymax=ucl, group=State, color=State) + 
                   geom_point() + geom_line() + geom_errorbar(width=0.05) + facet_grid(name ~ .) + 
                   scale_y_log10() + ylab("Estimated Number of Fires")
```

The table below lists the Bayesian estimates for the number of fires for each
year from 2019 to 2023. The 'Total' lines represent an estimate of the total 
number of such fires across the entire history. Again, the 'Plugin' category
includes both EVs and plugin hybrids. 

```{r tesla-table}
n.est %>% filter(is.na(Year) | Year > 2018) %>% 
          arrange(name, Year) %>% select(name, State, Year, est, lcl, ucl) %>% 
          mutate(Year=as.character(Year), Year=ifelse(is.na(Year), "Total", Year)) %>% 
          knitr::kable()
```

Similarly, estimating the rate at which such fires are increasing, we get 
the following table and figure.

```{r print-slope}
slp <- 
  stan.slope %>% sapply(function(x) extract(x, "r")[[1]] ) %>% exp() %>% subtract(1) %>% 
                 apply(2, function(x) c(mean(x), quantile(x, c(0.025, 0.975)))) %>% 
                 t() %>% as.data.frame() %>% 
                 cbind(names(stan.slope) %>% strsplit(":") %>% do.call("rbind", .), .) %>% 
                 setNames(c("State", "Type", "slope", "lcl", "ucl"))

row.names(slp) <- NULL
slp %>% knitr::kable(digits=2)
for(i in 3:5) slp[[i]] <- slp[[i]] * 100
slp$pm <- with(slp, (ucl - lcl) / 2)
```

```{r fires-slope-fig}
figs$ev.slope <- n.est %>% filter(! is.na(est) & ! is.na(Year), name != "Tesla") %>% 
                ggplot() + aes(x=Year, y=est, ymin=lcl, ymax=ucl, group=State, color=State, fill=State) + 
                   geom_point() + geom_errorbar(width=0.05) + 
#                   geom_line() + geom_ribbon(alpha=0.3) +  
#                   geom_point(aes(y=slope), color="blue") + 
                   geom_line(aes(y=slope)) +
                   geom_ribbon(aes(ymin=s.lcl, ymax=s.ucl), alpha=0.3) + 
                   facet_grid(name ~ .) + 
                   ylab("Estimated Number of Fires")
```


Fires in Teslas are growing at a rate of about `r slp$slope[7] %>% round(0)`&nbsp;% 
(&pm; `r slp$pm[7] %>% round(1)`&nbsp;%) per year.
Fires in EVs in general and Plugin vehicles (including EVs) are growing even faster,
 likely reflecting the fact that non-Teslas are a growing proportion of the market.
 Fires in EVs in general are growing at a rate of about `r slp$slope[8] %>% round(0)`&nbsp;% 
 (&pm; `r slp$pm[8] %>% round(1)`&nbsp;%) per year, and in Plugin vehicles at a rate
 of about `r slp$slope[9] %>% round(0)`&nbsp;% (&pm; `r slp$pm[9] %>% round(1)`&nbsp;%) per
 year.

```{r plan-c}
nfpa <- nfpa %>% mutate(ratio=NFPA / NFIRS, vin.ratio=NFPA/VINs)
```

It is interesting that about 1 in `r nfpa$vin.ratio %>% mean(na.rm=T) %>% round(1)` 
car fires where the VIN is recorded generally end up in NFIRS, but only about 1 in 
`r with(results$Summary %>% filter(is.na(Year)), Tesla[2] / Both7[2]) %>% round(1)` 
EV fires where the VIN is recorded end up in NFIRS. That makes the NFPA estimates 
not terribly useful.

I don't understand the reason for the difference, but it probably needs to be 
discussed. One possible explanation is that the fire is much more likely to 
destroy things like VIN numbers for the for EV fires, or the aggressive EV fire
simply makes collecting that information too hazardous.

Florida reports VIN numbers for better than 1 in 2 EV fires in the 
state (based on these results), which is approximately the rate at which they report 
VIN numbers of all vehicle fires. That suggests they have had only about 36 EV
fires ever in Florida. This seems unrealistically low, especially when compared
to the rest of the country. Excluding Florida, VIN numbers are reported for only 
about 1 in 30 EV fires. 

This has a significant effect on estimates of the number of fires. Without Florida,
estimates are about 1.5 times larger than when Florida is included. Given that 
the Florida number seem unrealistically low, and the almost unreasonably high rate
at which this data suggest that Florida records VIN numbers for Li ion battery 
vehicle fires, the fire estimates made excluding Florida are likely more reliable.


```{r cleanup-01a}
# Save the relevant file, as below. Then delete everything if you are working in bookdown.
# That means that a `bookdown` flag needs to be set somehow. And then unset at the end.
save(list=intersect(keep2, ls()), file=paste0(intmdt_dir, "CarFires.RData"))
rm(list=setdiff(ls(), keep))
```
