---
title: "Batteries"
subtitle: "Car Fires"
output: html_document
date: "2025-07-25"
---

# Cars Fires

## Action Taken

```{r setup-01a, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)

library(magrittr)
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(rstan)
library(RPostgreSQL)

keep <- c(ls(), "figs")

if(! exists("csv_dir")){
  csv_dir <- "csv/"
}
if(! exists("intmdt_dir")){
  intmdt_dir <- "rdata/"
}

# Either load the completed data (in CarFires.RData) 
# or the relevant base data in CarFires.Data.RData.
ff <- list.files(intmdt_dir) %>% grep("RData", ., ignore.case=T, value=T)
if("CarFires.rdata" %in% tolower(ff)){
  fg <- "CarFires.rdata" %>% match(tolower(ff)) %>% { ff[.] }
} else if("CarFires.data.rdata" %in% tolower(ff)){
  fg <- "CarFires.data.rdata" %>% match(tolower(ff)) %>% { ff[.] }
} else {
  stop("The data files needed for the Car Fires analysis are not present!")
}
load(paste0(intmdt_dir, fg))
rm(ff, fg)

keep2 <- c("car.fires", "tesla", "test.cases", "results", "totals", "n.est", "car.actions",
           "nfpa", "stan.model", "stan.mode1", "stan.tesla", "stan.slope", "figs")
```

This looks at action taken for Li ion vehicle fires to see if there is anything 
useful there.

```{r tesla-action, eval=FALSE}
action.codes <- read.csv(paste0(csv_dir, "act_tak_codes.csv"))[,-1]
car.actions[,11:13] %>% as.matrix() %>% as.vector() %>% {.[! is.na(.) & . != '' & . != 'NULL']} %>%
           data.frame(code=.) %>% left_join(action.codes, by=join_by(code==code_value)) %>%
           with(table(code_descr)) %>% sort(decreasing=T)
```

Action taken does not give anything useful. Basically, the only thing it says is
"Put out fire" as the action taken. Not useful.

## Fire Estimates

```{r start-data}
################################################################################
########### This Section Prepares Data and Models                    ###########
################################################################################
```


```{r tesla-fires, eval=! exists("tesla.fires")}
#The data used here are found at this site:
# "https://www.tesla-fire.com/"
# 
# The data was downloaded and imported for use.

# tesla <- openxlsx2::wb_to_df("../Batteries/csv/Tesla Fire .xlsx", 1) %>% 
tesla <- paste0(csv_dir, "Tesla Fires.xlsx") %>% openxlsx2::wb_to_df(1) %>% 
         { setNames(., c("ID", "Year", "Date", "Country", "State", "Model", "Fatalities", "Description", 
                         paste0("source", 1:(ncol(.) - 8)))) } %>% 
         filter(! is.na(Year)) %>% 
         select(ID, Year, Date, Country, State, Model, Fatalities, Description) %>%
         mutate(Date=lubridate::ymd(Date))
```


```{r nfpa-fires, eval=! exists("nfpa")}
#The data used here is downloaded from the NFPA
# 
# The data was downloaded and imported for use.
nfpa <- openxlsx2::read_xlsx("csv/new_sales.xlsx", 4)
nfpa <- car.summ %>% full_join(nfpa %>% select(Year, Fires, Deaths, Injuries),
                       by=join_by(year==Year)) %>% 
             setNames(c("Year", "NFIRS", "VINs", "NFPA", "Deaths", "Injuries"))
rm(car.summ)
```

```{r vin-lookup, eval=! exists("test.cases")}
car.fires$mob_vin_no <- car.fires$mob_vin_no %>% str_replace_all("I", "1")
car.fires$mob_vin_no <- car.fires$mob_vin_no %>% str_replace_all("O", "0")

reg.check.vin <- data.frame(vin=str_pad(toupper(car.fires$mob_vin_no), 17, "right", "*")) %>% 
                      mutate(r0=str_detect(vin, "[ -]|^[0-9*]+$", negate=T),
                             r1=str_detect(vin, "^[0-9A-HJ-NPR-Z*]{12}"),
                             r2=str_detect(vin, "[0-9*]{5}$"),
                             r3=str_detect(vin, "^[0-9A-Z*]{8}([0-9X])"),
                             r4=str_detect(vin, "^[0-9A-Z*]{9}([^0Z])"))
# Check digit in VIN
coded <- data.frame(text=c(0:9, LETTERS, "*"),
                    value=c(0:9, 1:9, 1:9, 2:9, NA))
coded$value[coded$text %in% c('I', 'O' , 'Q')] <- NA
multiplies <- c(8:2, 10, 0, 9:2)

tmp <- str_split_fixed(reg.check.vin$vin, "", 17)
tmp <- coded$value[match(tmp, coded$text)] %>% matrix(ncol=17) %>% multiply_by_matrix(multiplies) %>%
       magrittr::mod(11)

reg.check.vin$r5 <- substring(reg.check.vin$vin, 9, 9) == c(0:9, 'X', '@')[ifelse(is.na(as.integer(tmp)), 11, as.integer(tmp)) + 1]

ndx <- with(reg.check.vin, r0 & r1 & r2 & r3 & r4)
test.cases <- data.frame(vin=str_pad(toupper(car.fires$mob_vin_no), 17, "right", "*")[ndx],
                         year=car.fires$mob_year[ndx],
                         CheckValid=reg.check.vin$r5[ndx]) %>% 
                  mutate(wmi=substring(vin, 1, 3),
                         wmi=ifelse(substring(wmi, 3, 3) == '9', paste0(wmi, substring(vin, 12, 14)), wmi),
                         rest=substring(vin, 4,17)) %>% unique()

dbExecute(con, "truncate \"tmpVinCases\"")
dbWriteTable(con, "tmpVinCases", test.cases, row.names=F, append=T)

out.cases <- dbGetQuery(con, "SELECT * FROM \"GetElements\"")

test.cases <- out.cases %>% 
       group_by(vin, ModelYear, InputYear, CheckValid, VehicleType, nMakes, Make0, Element) %>% 
       summarize(value=first(Value)) %>%
       pivot_wider(id_cols=1:7, names_from=Element, values_from=value) %>% 
       mutate(`NA`=NULL, 
              Engine=`Engine Model`, 
              Fuel1=`Fuel Type - Primary`,
              Fuel2=`Fuel Type - Secondary`,
              Electric=`Electrification Level`,
              Battery=`Battery Type`,
              .keep="unused")

# test.cases$li.battery <- (! is.na(test.cases$Fuel1) & test.cases$Fuel1 == 'Electric') | 
#                          (! is.na(test.cases$Fuel2) & test.cases$Fuel2 == 'Electric')

car.fires$Tesla <- str_detect(car.fires$mob_vin_no, '^(5YJ|7SA)')
car.fires$BEV   <- car.fires$mob_vin_no %in% test.cases$vin[str_detect(test.cases$Electric, 'BEV')]
car.fires$Plugin<- car.fires$mob_vin_no %in% test.cases$vin[str_detect(test.cases$Electric, 'BEV|PHEV')]
car.fires$Hybrid<- car.fires$mob_vin_no %in% test.cases$vin[str_detect(test.cases$Electric, 'BEV|HEV')]
```


```{stan v2, output.var="stan.model", eval=! exists("stan.model")}
data {
  int<lower=0>       S1;         // Number of cars in sample 1
  int<lower=0>       S2;         // Number of cars in sample 2
  int<lower=0>       S12;        // Number of cars in both samples
  real               alpha;
  real               scale_p;
}

transformed data {
  int      which;
  real     theta;
  real     min_N;
  real     scale_N;

  if(S1 <= S2){
    which = 1;
    theta = alpha ^ (1. / S2);
  } else {
    which = 2;
    theta = alpha ^ (1. / S1);
  }
  if(S12 == 0){
    scale_N = sqrt(S1 * S2 / ((1 - alpha ^ (1. / S1)) * (1 - alpha ^ (1. / S2))));
  } else {
    scale_N = S1 * S2 / S12;
  }
  min_N = max(S1, S2) / scale_N;
}

parameters {
  vector<lower=0, upper=1>[2]   p;
  real<lower=min_N>             N_hat;
}

transformed parameters {
  real    N = N_hat * scale_N;
}

model {

  target += lgamma(N + 1) - lgamma(S1 + 1) - lgamma(N - S1 + 1)
            + S1 * log(p[1]) + (N - S1) * log(1 - p[1]);
  target += lgamma(N + 1) - lgamma(S2 + 1) - lgamma(N - S2 + 1)
            + S2 * log(p[2]) + (N - S2) * log(1 - p[2]);
  if(S12 == 0){
    theta ~ beta((1 - p[which]) * scale_p, p[which] * scale_p);
  } else {
    target += lgamma(S1 + 1) - lgamma(S12 + 1) - lgamma(S1 - S12 + 1)
              + S12 * log(p[2]) + (S1 - S12) * log(1 - p[2]);
  }
}

generated quantities {
}
```

```{stan stan-slope-v, output.var="stan.mode1", eval=! exists("stan.mode1")}
// This differs from the version above in that it computes a slope over time
// instead of computing estimated values for each year separately.
data {
  int<lower=1>          T;       // Number of time periods
  vector[T]             S1;      // Sample 1 Counts
  vector[T]             S2;      // Sample 2 Counts
  vector[T]             So;      // Sample Overlap Counts
  int<lower=0, upper=1> slope1;  // 0: No; 1: Yes
  int<lower=0, upper=1> slope2;  // 0: No; 1: Yes
}

transformed data {

  matrix[T,3]    dats = append_col(append_col(S1, S2), So);
  
  real     min_N;
  real     scale_N = 0;
  int      k=0;
  for(i in 1:T){
    if(So[i] > 0){
      scale_N += S1[i] * S2[i] / So[i];
      k = k + 1;
    }
  }
  scale_N = scale_N / k;
  
  min_N = max(append_row(S1, S2)) / scale_N;
}

parameters {
// The affine transformations make sure that Stan's initial sampling occurs
// near the actual expected values. Otherwise bad initial values become too
// common to reliably come up with results.
  vector<offset=-2>[1 + slope1]  p1_lgt;
  vector<offset=-2>[1 + slope2]  p2_lgt;
  real<multiplier=0.25>          r;
  real<lower=min_N>              N_hat;
}

transformed parameters {
  vector[T]    N;
  vector[T]    p1;
  vector[T]    p2;
  for(i in 1:T){
    N[i] = N_hat * scale_N * exp((i - 1) * r);
    if(slope1 == 1){
      p1[i] = inv_logit(p1_lgt[1] + (i - 1) * p1_lgt[2]);
    } else {
      p1[i] = inv_logit(p1_lgt[1]);
    }
    if(slope2 == 1){
      p2[i] = inv_logit(p2_lgt[1] + (i - 1) * p2_lgt[2]);
    } else {
      p2[i] = inv_logit(p2_lgt[1]);
    }
  }
}

model {
  matrix[T, 3]   p_hat = append_col(append_col(p1, p2), p1 .* p2);
  for(i in 1:T){
    for(j in 1:3){
      target += lgamma(N[i] + 1) - lgamma(dats[i,j] + 1) - lgamma(N[i] - dats[i,j] + 1)
              + dats[i,j] * log(p_hat[i,j]) + (N[i] - dats[i,j]) * log(1 - p_hat[i,j]);
  }}
}

generated quantities {
}
```

```{r start-results}
################################################################################
########### Results Start Here                                       ###########
################################################################################
```


```{r car-fire-est}
# This section just does the basic Capture-Recapture calculation
# (after building the data tables, of course).
results <- list(Fires=tesla %>% mutate(year=Year) %>% filter(tolower(Country) == "usa") %>% 
                                group_by(year) %>% summarize(Tesla=n()) %>% 
        full_join(car.fires %>% filter(Tesla) %>% group_by(year) %>% summarize(t.fires=n())) %>% 
        full_join(car.fires %>% filter(BEV, year > 2004) %>% group_by(year) %>% summarize(BEV=n())) %>% 
        full_join(car.fires %>% filter(Plugin, year > 2004) %>% group_by(year) %>% summarize(Plugin=n())) %>% 
        full_join(car.fires %>% filter(Hybrid, year > 2004) %>% group_by(year) %>% summarize(Hybrid=n())) %>% 
        arrange(year),
                both0=tesla %>% filter(tolower(Country) == 'usa') %>% 
                      inner_join(car.fires %>% filter(Tesla),
                                 by=join_by(Date == inc_date, State==state)) %>% 
                      select(Year, Date, State, Model, Fatalities, 
                             mob_model, mob_year, mob_vin_no, Description) %>% unique(),
                both4=tesla %>% filter(tolower(Country) == 'usa') %>% 
                      inner_join(car.fires %>% filter(Tesla) %>% 
                                   mutate(ldate=inc_date - 4, hdate=inc_date + 4),
                                 by=join_by(Date >= ldate, Date <= hdate, State==state)) %>% 
                      select(Year, Date, inc_date, State, Model, Fatalities, 
                             mob_model, mob_year, mob_vin_no, Description) %>% unique(),
                both7=tesla %>% filter(tolower(Country) == 'usa') %>% 
                      inner_join(car.fires %>% filter(Tesla) %>% 
                                   mutate(ldate=inc_date - 7, hdate=inc_date + 7),
                                 by=join_by(Date >= ldate, Date <= hdate, State==state)) %>% 
                      select(Year, Date, inc_date, State, Model, Fatalities, 
                             mob_model, mob_year, mob_vin_no, Description) %>% unique()
)

results$Summary <- 
  results$Fire %>% full_join(results$Both0 %>% group_by(Year) %>% summarize(both0=n()), 
                             by=join_by(year == Year)) %>% 
                   full_join(results$Both4 %>% group_by(Year) %>% summarize(both4=n()), 
                             by=join_by(year == Year)) %>% 
                   full_join(results$Both7 %>% group_by(Year) %>% summarize(both7=n()), 
                             by=join_by(year == Year)) %>% 
              filter(year > 2005, year < 2024) %>% 
              { rbind(., c(NA, colSums(.[,2:9], na.rm=T))) } %>% 
              rename(Year = year) %>% 
              mutate(Year = as.integer(Year), 
#                     tesla0  = Tesla * t.fires / both0,
                     tesla4  = Tesla * t.fires / both4,
#                     tesla7  = Tesla * t.fires / both7,
#                     bev0    = Tesla * BEV / both0,
                     bev4    = Tesla * BEV / both4,
#                     bev7    = Tesla * BEV / both7,
#                     plugin0 = Tesla * Plugin / both0,
                     plugin4 = Tesla * Plugin / both4,
#                     plugin7 = Tesla * Plugin / both7
                     ) %>% arrange(Year)

results$Summary[,-c(7,9)] %>% knitr::kable(digits=1)
```

The table above list capture-recapture estimate for the number of Tesla, EV, 
and Plugin-Hybrid vehicle fires. Note that the last column (the 'plugin4' column)
represents the total of EV and plugin-hybrid fires. The final line (with no 
year) represents the estimated total for *all* years in the sample.

```{r run_stan1, eval=! exists("stan.tesla")}
n.est <- results$Summary %>%
         select(Year, Tesla, t.fires, BEV, Plugin, both4) %>% 
         pivot_longer(c(t.fires, BEV, Plugin)) %>% 
         mutate(Tesla=as.integer(Tesla),
                both4=as.integer(both4),
                name=ifelse(name == 't.fires', "Tesla", name),
                value=as.integer(value),
                est=as.numeric(NA),
                lcl=est,
                ucl=est,
                med=est)
stan.tesla <- list()
for(i in 1:nrow(n.est)){
  if(is.na(n.est$Year[i]) | n.est$Year[i] > 2018){
    nm <- paste0(sub("t.fires", "Tesla", n.est$name[i]), ":", 
                 ifelse(is.na(n.est$Year[i]), "Total", n.est$Year[i]))
    st.data <- with(n.est, list(S1=Tesla[i], S2=value[i], 
                                S12=ifelse(is.na(both4[i]), 0, both4[i]),
                                alpha=0.1, scale_p=10))
    stan.tesla[[nm]] <- sampling(stan.model, data=st.data, refresh=0)
    n.est[i,6:9] <- stan.tesla[[nm]] %>% extract("N") %>% magrittr::extract2(1) %>% 
                    {c(mean(.), quantile(., c(0.025, 0.975, 0.5)))} %>% t()
  }
}
```

```{r run_stan2, eval=! exists("stan.slope")}
st.data <- list(T=2023 - 2018, 
                S1=results$Summary$Tesla[14:18], 
                S2=results$Summary$t.fires[14:18], 
                So=ifelse(is.na(results$Summary$both4[14:18]), 0, results$Summary$both4[14:18]),
                slope1=0, 
                slope2=0)

st.data$S2 <- results$Summary$t.fires[14:18]
stan.slope <- list(Tesla=sampling(stan.mode1, data=st.data, refresh=0))
st.data$S2 <- results$Summary$BEV[14:18]
stan.slope$BEV <- sampling(stan.mode1, data=st.data, refresh=0)
st.data$S2 <- results$Summary$Plugin[14:18]
stan.slope$Plugin <- sampling(stan.mode1, data=st.data, refresh=0)

```


```{r more-n-est}
# Add the slope into the n.est data frame
n.est <- stan.slope %>% lapply(function(y) y %>% extract("N") %>% magrittr::extract2(1) %>% 
                                  apply(2, function(x) c(mean(x), quantile(x, c(0.025, 0.975, 0.5)))) %>% 
                                  t() ) %>% 
               do.call("rbind", .) %>% 
               as.data.frame() %>% setNames(c("slope", "s.lcl", "s.ucl", "s.med")) %>%
               cbind(name=rep(c("Tesla", "BEV", "Plugin"), each=5), 
                     Year=rep(2019:2023, 3), .) %>% 
               left_join(n.est, .)
```

```{r fires-fig}
if(! exists("figs")) figs <- list()
figs$ev <- n.est %>% filter(! is.na(est) & ! is.na(Year)) %>% 
                ggplot() + aes(x=Year, y=est, ymin=lcl, ymax=ucl) + geom_point() + geom_line() +
                   geom_ribbon(alpha=0.3) + facet_grid(name ~ .) + 
                   ylab("Estimated Number of Fires")
```

The table below lists the Bayesian estimates for the number of fires for each
year from 2019 to 2023. The 'Total' lines represent an estimate of the total 
number of such fires across the entire history. Again, the 'Plugin' category
includes both EVs and plugin hybrids. 

```{r tesla-table}
n.est %>% filter(is.na(Year) | Year > 2018) %>% 
          arrange(name, Year) %>% select(name, Year, est, lcl, ucl) %>% 
          mutate(Year=as.character(Year), Year=ifelse(is.na(Year), "Total", Year)) %>% 
          knitr::kable()
```

Similarly, estimating the rate at which such fires are increasing, we get 
the following table and figure.

```{r print-slope}
stan.slope %>% sapply(function(x) extract(x, "r")[[1]] ) %>% exp() %>% subtract(1) %>% 
               apply(2, function(x) c(mean(x), quantile(x, c(0.025, 0.975)))) %>% 
               t() %>% as.data.frame() %>% setNames(c("slope", "lcl", "ucl")) %>% 
               knitr::kable(digits=2)
```

```{r fires-slope-fig}
figs$ev.slope <- n.est %>% filter(! is.na(est) & ! is.na(Year)) %>% 
                ggplot() + aes(x=Year, y=est, ymin=lcl, ymax=ucl) + geom_point() + 
#                   geom_line() + geom_ribbon(alpha=0.3) +  
#                   geom_point(aes(y=slope), color="blue") + 
                   geom_line(aes(y=slope), color="blue") +
                   geom_ribbon(aes(ymin=s.lcl, ymax=s.ucl), alpha=0.3, color="blue") + 
                   facet_grid(name ~ .) + 
                   ylab("Estimated Number of Fires")
```


Fires in Teslas are growing at a rate of about 39&nbsp;% (&pm; 12.5&nbsp;%) per year.
Fires in EVs in general and Plugin vehicles (including EVs) are growing even faster,
 likely reflecting the fact that non-Teslas are a growing proportion of the market.
 Fires in EVs in general are growing at a rate of about 44&nbsp;% (&pm; 12.2&nbsp;%) per
 year, and in Plugin vehicles at a rate of about 42&nbsp;% (&pm; 9.7&nbsp;%) per
 year.

```{r plan-c}
nfpa <- nfpa %>% mutate(ratio=NFPA / NFIRS, vin.ratio=NFPA/VINs)
```

It is interesting that about 1 in 3 car fires where the VIN is recorded generally 
end up in NFIRS, but only about 1 in 6.5 EV fires where the VIN is recorded end up 
in NFIRS. That makes the NFPA estimates not terribly useful.

I don't understand the reason for the difference, but it probably needs to be 
discussed. One possible explanation is that the fire is much more likely to 
destroy things like VIN numbers for the for EV fires, or the aggressive EV fire
simply makes collecting that information too hazardous.



```{r cleanup-01a}
# Save the relevant file, as below. Then delete everything if you are working in bookdown.
# That means that a `bookdown` flag needs to be set somehow. And then unset at the end.
save(list=intersect(keep2, ls()), file=paste0(intmdt_dir, "CarFires.RData"))
rm(list=setdiff(ls(), keep))
```
