---
title: "Batteries"
subtitle: "NFIRS Data"
output: html_document
date: "2024-12-13"
---

# NFIRS

```{r setup-02, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)

library(magrittr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(rstan)

keep <- c(ls(), "figs")

if(! exists("csv_dir")){
  csv_dir <- "csv/"
}
if(! exists("intmdt_dir")){
  intmdt_dir <- "rdata/"
}

source("load.files.R")
if(
  load.data(final=c("NFIRS.RData"), 
            data =c("nfirs.data.RData"), 
            dir=intmdt_dir) == 1
  ) stop("The data files needed for the NFIRS analysis are not present!")

rm(load.data)

keep2 <- list(NFIRS=c("data.summ", "details", "selected", "structures", 
                      "stan.summ", "stan.summ.out", "tod.results", "seasonality", 
                      "sev.results", "sev.base", "room.res", "figs")
              )
if(! exists("figs")) figs <- list()
```

## Overall Statistics

```{r data-summary}
figs$figure.00 <- 
    data.summ %>% filter(year > 2010, inc_type == '11') %>% 
      mutate(ratio.battery=battery_casualties / battery_fires, 
             ratio.structure=casualties/fires) %>% 
      select(c("year", starts_with("ratio"))) %>% 
      pivot_longer(2:3, names_prefix="ratio[.]", values_to="ratio") %>% 
      ggplot() + aes(x=year, y=ratio, group=name, color=name) + 
          geom_point() + geom_line() + expand_limits(y=0) + 
          scale_x_continuous(breaks=2010 + 0:3 * 5, minor_breaks=2010:2025)
print(figs$figure.00)
```

```{stan summ, output.var="stan.summ", eval=! exists("stan.summ")}
data {
  int           fires;
  int           casualties;
  int           batt_fire;
  int           batt_cas;
}

transformed data {
}

parameters {
  real<lower=0, upper=1>   p_batt;
  real<lower=0, upper=1>   p_cas;
  real<lower=0, upper=1>   p_batt_cas;
}

transformed parameters {
}

model {
  casualties ~ binomial(fires, p_cas);
  batt_fire  ~ binomial(fires, p_batt);
  batt_cas ~ binomial(batt_fire, p_batt_cas);
}

generated quantities {
  real            r_batt_cas = p_batt_cas / p_cas;
}
```


```{r summ-table}
tmp <- data.summ %>% filter(inc_type == 11) %>% group_by(cut(year, c(-Inf, 2004, 2010, 2017, Inf))) %>% 
  summarize(fires=sum(fires), cas=sum(casualties), batt_fires=sum(battery_fires),
            batt_cas=sum(battery_casualties)) %>% 
  mutate(fires=rev(cumsum(rev(fires))), cas=rev(cumsum(rev(cas))), 
         batt_fires=rev(cumsum(rev(batt_fires))), batt_cas=rev(cumsum(rev(batt_cas)))) %>% 
  magrittr::extract(3, 2:5) %>% unlist() %>% matrix(ncol=2) %>% {rbind(., .[2,] / .[1,])} %>%
  as.data.frame() %>% cbind(type=c("Fires", "Casualties", "Ratio"), .) %>% 
  setNames(c("type", "Structure", "Battery")) %>% 
  mutate(Ratio = Battery / Structure)

stan.summ.out <- tmp %>% with(list(fires=Structure[1], casualties=Structure[2], 
                                   batt_fire=Battery[1], batt_cas=Battery[2])) %>%
  sampling(stan.summ, data=., refresh=F)

tmp2 <- stan.summ.out %>% rstan::extract() %>% 
        sapply(function(x) c(mean(x), sd(x), quantile(x, c(0.025, 0.975)))) %>% as.data.frame() %>%
        cbind(value=c("mean", "sd", "lcl", "ucl"), .)
row.names(tmp2) <- NULL
```

This table shows summary information on fires, battery fires, and injuries.

```{r summary-a}
tmp %>% knitr::kable()
```

And this table shows confidence intervals on all the ratios in the table above.
Ratios and confidence intervals for this table are computed in Stan. 
What is significant here is that injuries from battery fires occur at almost 
double the rate of regular structure fires (`r tmp2$r_batt_cas[1] %>% formatC(digits=2, format="f")`, 
Confidence interval: `r tmp2$r_batt_cas[3] %>% formatC(digits=2, format="f")` - 
`r tmp2$r_batt_cas[4] %>% formatC(digits=2, format="f")`).

```{r summary-b}
tmp2[,1:5] %>% knitr::kable()
rm(tmp, tmp2)
```

## Selected (large) departments

Data are from NFIRS. Initially we look at summaries of data for selected 
departments to see what the data looks like.

```{r selected-fig}
selected %>% mutate(fires=total - not_fire) %>% 
             select(state, fdid, fd_name, year, type, fires) %>% 
             pivot_wider(names_from=type, values_from=fires) %>% 
             filter(year > 2018) %>% group_by(fd_name) %>% 
             summarize(Total=sum(Other, na.rm=T), Battery=sum(Battery, na.rm=T)) %>% 
             mutate(Total=Total + Battery, ratio=Total / Battery) %>% 
             arrange(desc(ratio)) %>% 
     knitr::kable(col.names=c("Department", "All Fires", "Battery Fires", "Ratio"))
```

Basically, what is tells me is that NFIRS is terrible when it comes to 
Li ion battery fires. The very wide range of number of battery fires per 
reported fire is not realistically representative of the real world, and can 
only mean that departments vary wildly in the extent to which they identify
battery fires. Worse, the largest departments are typically the worst at it.

## Time of Day

```{r tod}
tod.results <- list(
  all= details %>% 
               with(aggregate(data.frame(fires=fdid), 
                              list(year=year - 2015, tod=factor(tod)), length)) %>%
               glm(fires ~ tod + year, data=., family=quasipoisson),
  home=details %>% filter(prop_use %in% c('419', '429')) %>%
               with(aggregate(data.frame(fires=fdid), 
                              list(year=year - 2015, tod=factor(tod)), length)) %>%
               glm(fires ~ tod + year, data=., family=quasipoisson),
  other=details %>% filter(! prop_use %in% c('419', '429')) %>%
               with(aggregate(data.frame(fires=fdid), 
                              list(year=year - 2015, tod=factor(tod)), length)) %>%
               glm(fires ~ tod + year, data=., family=quasipoisson))

tmp.df <- NULL

for(i in names(tod.results)){
  uuu <- tod.results[[i]]
  ndx <- uuu %>% coef() %>% { grep("tod", names(.)) }
  tmp.df <- rbind(tmp.df, 
                  data.frame(group=i,
                       tod=0:24, 
                       val=c(0, uuu %>% coef() %>% { .[ndx] }, 0),
                       se =c(0, uuu %>% vcov() %>% diag() %>% {.[ndx]} %>% sqrt(), 0)))
}

tmp.df$lcl <- with(tmp.df, val + qnorm(0.025) * se)
tmp.df$ucl <- with(tmp.df, val + qnorm(0.975) * se)

rm(uuu, i, ndx)

figs$figure.07 <- tmp.df %>% ggplot() + aes(x=tod, y=val, ymin=lcl, ymax=ucl) +
           geom_point() + geom_line() + geom_ribbon(alpha=0.3) + facet_wrap(~ group) +
           scale_x_continuous(breaks=0:4 * 6, minor_breaks=0:24) + theme_bw()

figs$figure.07
```

Home fires are about 2/3 of the total battery fires in NFIRS. The figure above
shows the effect of time of day on battery fires for all battery fires, home 
battery fires, and battery fires not in homes. All follow a fairly typical 
day-night cycle where battery fires are more common during the day than at night.

Home fires vary from that slightly in that they are more common when people are
active at home. So there is a slight peak in the number of fires between 7 and 9
am, and a much larger peak in the late afternoon or early evening. Non-home 
battery fires have a relatively stable fire rate between the hours of 8am and 
8pm and tend to have a fire rate that is much lower outside of those hours.

## Seasonality

```{r season}
seasonality <- list(month=details %>% filter(prop_use %in% c('419', '429')) %>% 
                      mutate(yr = year - 2015, month=factor(month)) %>% 
                      group_by(yr, month) %>% summarize(fires=n()) %>%
                      glm(fires ~ 0 + yr + month, data=., family=quasipoisson),
                    seas=details %>% filter(prop_use %in% c('419', '429')) %>% 
                      mutate(yr = year - 2015, season=factor(month %/% 3 %>% mod(4))) %>% 
                      group_by(yr, season) %>% summarize(fires=n()) %>%
                      glm(fires ~ 0 + yr + season, data=., family=quasipoisson))

season.df <- summary(seasonality$month)$coefficients %>% as_tibble(rownames="var") %>% as.data.frame() %>%
             magrittr::extract(-1,)
season.df$var <- 1:12
# season.df[1, 2:5] <- c(0, 0, NA, NA)
season.df$var <- factor(season.df$var, labels=month.name)
season.df$lcl <- with(season.df, Estimate + `Std. Error` * qnorm(0.025))
season.df$ucl <- with(season.df, Estimate + `Std. Error` * qnorm(0.975))

tmp <- summary(seasonality$seas)$coefficients %>% as_tibble(rownames="var") %>% as.data.frame() %>%
             magrittr::extract(-1,)
# tmp[1, 2:5] <- c(0, 0, NA, NA)
tmp$var <- 0:3
tmp$lcl <- with(tmp, Estimate + `Std. Error` * qnorm(0.025))
tmp$ucl <- with(tmp, Estimate + `Std. Error` * qnorm(0.975))
names(tmp) <- c("var.a", "Season", "Season.sd", "Season.t", "Season.p", "Season.lcl", "Season.ucl")
uu <- as.integer(season.df$var) %/% 3 %>% mod(4) %>% match(0:3)

season.df$season <- tmp$Season[uu]
season.df$seas.lcl <- tmp$Season.lcl[uu]
season.df$seas.ucl <- tmp$Season.ucl[uu]


figs$figure.08 <- season.df %>% ggplot() + aes(x=var, y=Estimate, ymin=lcl, ymax=ucl, group="Month", color="Month") + 
                 geom_point() + geom_line() + geom_ribbon(alpha=0.3) + 
                 geom_line(aes(y=season, group="Season", color="Season")) +
                 geom_ribbon(aes(ymin=seas.lcl, ymax=seas.ucl, group="Season", color="Season"), alpha=0.3) + 
                 theme_bw()

figs$figure08a <- season.df %>% ggplot() + aes(x=var, y=Estimate, ymin=lcl, ymax=ucl, group="Month", color="Month") + 
                 geom_point() + geom_errorbar(width=0.1) + 
                 geom_step(aes(y=season, group="Season", color="Season"), position=position_nudge(x=-.5)) +
                 geom_ribbon(aes(ymin=seas.lcl, ymax=seas.ucl, group="Season", color="Season"), alpha=0.3) + 
                 theme_bw()

figs$figure.08
```

Looking at season, there is a slight peak in home battery fires over the summer.
Beyond that, it is possible that December battery fires are anomalously high, but 
such results are not statistically significant.

## Severity

```{r sev-base}
sev.base <- list(room=data.summ %>% filter(inc_type == '11', year > 2010) %>% select(-(1:3)) %>% 
                   glm(cbind(room, fires - room)~ 1, data=., family=quasibinomial),
                 struc=data.summ %>% filter(inc_type == '11', year > 2010) %>% select(-(1:3)) %>% 
                   glm(cbind(struc, fires - struc)~ 1, data=., family=quasibinomial),
                 beyond=data.summ %>% filter(inc_type == '11', year > 2010) %>% select(-(1:3)) %>% 
                   glm(cbind(beyond, fires - beyond)~ 1, data=., family=quasibinomial))

sev.base$summary <- sapply(sev.base, 
                           function(x){
                             q <- coef(x)
                             s <- sqrt(vcov(x)[1,1])
                             c(q, s, plogis(q), 
                               plogis(q + qnorm(0.025) * s), 
                               plogis(q + qnorm(0.975) * s))
                             }
                           ) %>% 
       t() %>% as.data.frame() %>% setNames(c("logit(p)", "s", "p", "lcl", "ucl"))
```

```{r sev}
sev.dat <- details %>% filter(! fire_sprd %in% c("", "NULL")) %>% 
           mutate(sev = factor(ifelse(fire_sprd == 3, 4, fire_sprd)))

sev.results <- list(
  all= sev.dat %>% mutate(yr = year - 2015) %>% group_by(yr, sev) %>% 
                   summarize(fires = n()) %>% arrange(desc(sev)) %>% 
                   mutate(batts=cumsum(fires), not=sum(fires) - batts) %>% 
                   filter(sev != '1') %>%
                   glm(cbind(batts, not) ~ sev, data=., family=quasibinomial),
  home=sev.dat %>% filter(prop_use %in% c('419', '429')) %>%
                   mutate(yr = year - 2015) %>% group_by(yr, sev) %>% 
                   summarize(fires = n()) %>% arrange(desc(sev)) %>% 
                   mutate(batts=cumsum(fires), not=sum(fires) - batts) %>% 
                   filter(sev != '1') %>%
                   glm(cbind(batts, not) ~ sev, data=., family=quasibinomial),
  other=sev.dat %>% filter(! prop_use %in% c('419', '429')) %>%
                   mutate(yr = year - 2015) %>% group_by(yr, sev) %>% 
                   summarize(fires = n()) %>% arrange(desc(sev)) %>% 
                   mutate(batts=cumsum(fires), not=sum(fires) - batts) %>% 
                   filter(sev != '1') %>%
                   glm(cbind(batts, not) ~ sev, data=., family=quasibinomial))

sev.df <- NULL

for(i in names(sev.results)){
  uuu <- sev.results[[i]]
  sev.df <- rbind(sev.df, 
                  data.frame(group=i,
                       sev=factor(c("room", "structure", "beyond"),
                                  levels=c("room", "structure", "beyond")), 
                       val=uuu %>% coef(),
                       se =uuu %>% vcov() %>% diag() %>% sqrt()))
}

sev.df$lcl <- with(sev.df, val + qnorm(0.025) * se)
sev.df$ucl <- with(sev.df, val + qnorm(0.975) * se)

rm(uuu, i, sev.dat)

figs$figure.09 <- sev.df %>% mutate(val=plogis(val), lcl=plogis(lcl), ucl=plogis(ucl))%>%
           ggplot() + aes(x=sev, y=val, ymin=lcl, ymax=ucl) +
           geom_point() + geom_path(group=1) + geom_ribbon(group=1, alpha=0.3) +
           facet_wrap(~ group) + theme_bw()

figs$figure.09a <- sev.df %>% filter(group != "all") %>% 
           mutate(val=plogis(val), lcl=plogis(lcl), ucl=plogis(ucl))%>%
           ggplot() + aes(x=sev, y=val, ymin=lcl, ymax=ucl, group=group, color=group) +
           geom_point() + geom_errorbar(width=.05) + 
           xlab("") + ylab("Probability") + theme_bw()


figs$figure.09b <- sev.base$summary[,-3] %>% 
   { cbind(group="baseline", sev=c("room", "structure", "beyond") , .)} %>% 
   mutate(lcl=qlogis(lcl), ucl=qlogis(ucl)) %>% 
   setNames(c("group", "sev", "val", "se", "lcl", "ucl")) %>% 
   rbind(sev.df, .) %>% filter(group != "all") %>% 
   mutate(val=plogis(val), lcl=plogis(lcl), ucl=plogis(ucl))%>%
   ggplot() + aes(x=sev, y=val, ymin=lcl, ymax=ucl, group=group, color=group) +
      geom_point() + geom_errorbar(width=.05) + 
      xlab("") + ylab("Probability") + theme_bw()

figs$figure.09
```

This represents the probability of a battery fire of the size specified or bigger 
for all battery fires, home battery fires, and all battery structure fires that 
are not in homes.

```{r fig6a}
figs$figure.7b
```

This specifically compares those values for home v. not-home fires. Home fires
are more likely to expand to the room of origin (beyond the object of origin)
than other fires. But other fires are more likely to expand beyond the room 
or origin or beyond the structure of origin. It seems likely this reflects
the size of the battery pack that initiated the blaze. Home fires are more 
likely to expand beyond the object of origin but less likely to get bigger 
than that.

## Room in Home

```{r room}
# The following table takes the area of origin field in `details` and 
# consolidates them into a smaller number of (hopefully) more meaningful
# categories.
area_orig <- data.frame(area_orig=sort(unique(details$area_orig)))
area_orig$NewCode <- c("Other", "Hall/Stair", "Exterior", "Hall/Stair", "Entrance/Exit",
                       "Entrance/Exit", "Common Area", "Common Area", "Nonresidential",
                       "Nonresidential", "Common Area", "Nonresidential", "Other",
                       "Bedroom", "Bedroom", "Dining Room", "Kitchen", "Bathroom",
                       "Laundry", "Office", "Bathroom", "Nonresidential", 
                       "Nonresidential", "Nonresidential", "Nonresidential", 
                       "Nonresidential", "Nonresidential", "Other", "Storage", 
                       "Storage", "Closet", "Storage", "Nonresidential", "Other", 
                       "Garage", "Other", "Equipment", "Equipment", "Nonresidential",
                       "Equipment", "Equipment", "Equipment", "Nonresidential", 
                       "Structure", "Structure", "Exterior", "Nonresidential", 
                       "Nonresidential", "Structure", "Exterior", "Nonresidential",
                       "Exterior", "Garage", "Garage", "Garage", "Garage", "Other",
                       "Garage", "Exterior", "Exterior", "Nonresidential", 
                       "Nonresidential", "Exterior", "Exterior", "Nonresidential",
                       "Undetermined")

room.res <- details %>% merge(area_orig, by="area_orig") %>% 
            filter(prop_use %in% c('419', '429')) %>% 
            mutate(yr = year - 2015, rooms=factor(NewCode) %>% relevel("Bedroom")) %>% 
            group_by(yr, rooms) %>% summarize(fires=n()) %>%
            glm(fires ~ yr + rooms, data=., family=quasipoisson)

room.df <- data.frame(room=room.res %>% coef() %>% names(),
                       val=room.res %>% coef(),
                        se=room.res %>% vcov() %>% diag() %>% sqrt())
room.df <- room.df[-2,]
room.df$room[1] <- "Bedroom"
room.df[1,2:3] <- 0
room.df$room <- room.df$room %>% sub("^rooms", "", .)
room.df$lcl <- with(room.df, val + qnorm(0.025) * se)
room.df$ucl <- with(room.df, val + qnorm(0.975) * se)

room.df$room <- factor(room.df$room, 
                       levels=room.df$room[order(room.df$val, decreasing=TRUE)])

figs$figure.0a <- room.df %>% ggplot() + aes(x=room, y=val, ymin=lcl, ymax=ucl) +
            geom_point() + geom_errorbar(width=0.1) + 
            scale_x_discrete(guide = guide_axis(angle = 90)) + xlab("") + ylab("")

figs$figure.0a
```

```{r play-with-rankings}
N <- 1000
mu <- coef(room.res)[-(1:2)]
Sigma <- vcov(room.res)[-(1:2), -(1:2)] %>% eigen()
nv <- length(mu)
nm <- c("Bathroom", names(mu) %>% sub("^rooms", "", .))

vals <- rnorm(nv * N) %>% matrix(ncol=N)
vals <- mu + Sigma$vectors %*% diag(sqrt(Sigma$values)) %*% vals
vals <- rbind(0, vals)

ranks <- as.integer(NA) %>% matrix(ncol=N, nrow=nv + 1)
for(i in 1:N){ ranks[,i] <- rank(vals[,i], ties.method="first") }
ranks <- ranks %>% t() %>% as.data.frame() %>% setNames(nm)

rank.list <- as.integer(0) %>% matrix(ncol=nv + 1, nrow=nv + 1)
colnames(rank.list) <- nm
row.names(rank.list) <- 1:(nv + 1)

for(i in names(ranks)){
  uuu <- ranks[[i]] %>% table() %>% { c(as.integer(names(.)), as.integer(.)) } %>% matrix(ncol=2)
  rank.list[uuu[,1], i] <- uuu[,2]
}

# I am probably not going to print this out, but I will keep it in case
# it ever becomes useful. The discussion below relies on this, though,
# for some of its conclusions.

rm(uuu, vals, ranks, i, nv, nm, N, mu, Sigma)
```

The garage is the number one area of origin for battery fires, followed by the 
bedroom and common areas. The kitchen is fourth, with other areas following after
those.

```{r cleanup-02}
# Save the relevant file, as below. Then delete everything if you are working in bookdown.
# That means that a `bookdown` flag needs to be set somehow. And then unset at the end.
for(i in names(keep2)){
  save(list=intersect(keep2[[i]], ls()), file=paste0(intmdt_dir, i, ".RData"))
}
rm(list=setdiff(ls(), keep))
```