---
title: "Batteries"
subtitle: "NFIRS Data"
output: html_document
date: "2024-12-13"
---

# NFIRS

```{r setup-02, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)

library(magrittr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(rstan)
library(loo)

keep <- c(ls(), "figs")

if(! exists("csv_dir")){
  csv_dir <- "csv/"
}
if(! exists("intmdt_dir")){
  intmdt_dir <- "rdata/"
}

# Either load the completed data (in NFIRS.RData) 
# or the relevant base data in NFIRS.Data.RData.
ff <- list.files(intmdt_dir) %>% grep("RData", ., ignore.case=T, value=T)
if("nfirs.rdata" %in% tolower(ff)){
  fg <- "nfirs.rdata" %>% match(tolower(ff)) %>% { ff[.] }
} else if("nfirs.data.rdata" %in% tolower(ff)){
  fg <- "nfirs.data.rdata" %>% match(tolower(ff)) %>% { ff[.] }
} else {
  stop("The data files needed for the NFIRS analysis are not present!")
}
load(paste0(intmdt_dir, fg))
rm(ff, fg)

# keep <- c(ls(), "figs")
keep2 <- c("data.summ", "tesla", "tesla_detail", "car.actions", "selected", "structures", "details",
           "stan.summ", "stan.summ.out", "struc.13", "cur.batt", "stan.model", "stan.out2", "stan.out3",
           "switchdate", "stan.model1", "st_out1", "stan.model2", "st_out2", "tod.results",
           "seasonality", "sev.results", "room.res", "loo_loo", "figs")

```

## Overall Statistics

```{r data-summary}
if(! exists("figs")) figs <- list()
figs$figure.00 <- 
    data.summ %>% filter(year > 2010, inc_type == '11') %>% 
      mutate(ratio.battery=battery_casualties / battery_fires, 
             ratio.structure=casualties/fires) %>% 
      select(c("year", starts_with("ratio"))) %>% 
      pivot_longer(2:3, names_prefix="ratio[.]", values_to="ratio") %>% 
      ggplot() + aes(x=year, y=ratio, group=name, color=name) + 
          geom_point() + geom_line() + expand_limits(y=0) + 
          scale_x_continuous(breaks=2010 + 0:3 * 5, minor_breaks=2010:2025)
print(figs$figure.00)
```

```{stan summ, output.var="stan.summ", eval=! exists("stan.summ")}
data {
  int           fires;
  int           casualties;
  int           batt_fire;
  int           batt_cas;
}

transformed data {
}

parameters {
  real<lower=0, upper=1>   p_batt;
  real<lower=0, upper=1>   p_cas;
  real<lower=0, upper=1>   p_batt_cas;
}

transformed parameters {
}

model {
  casualties ~ binomial(fires, p_cas);
  batt_fire  ~ binomial(fires, p_batt);
  batt_cas ~ binomial(batt_fire, p_batt_cas);
}

generated quantities {
  real            r_batt_cas = p_batt_cas / p_cas;
}
```


```{r summ-table}
tmp <- data.summ %>% filter(inc_type == 11) %>% group_by(cut(year, c(-Inf, 2004, 2010, 2017, Inf))) %>% 
  summarize(fires=sum(fires), cas=sum(casualties), batt_fires=sum(battery_fires),
            batt_cas=sum(battery_casualties)) %>% 
  mutate(fires=rev(cumsum(rev(fires))), cas=rev(cumsum(rev(cas))), 
         batt_fires=rev(cumsum(rev(batt_fires))), batt_cas=rev(cumsum(rev(batt_cas)))) %>% 
  magrittr::extract(3, 2:5) %>% unlist() %>% matrix(ncol=2) %>% {rbind(., .[2,] / .[1,])} %>%
  as.data.frame() %>% cbind(type=c("Fires", "Casualties", "Ratio"), .) %>% 
  setNames(c("type", "Structure", "Battery")) %>% 
  mutate(Ratio = Battery / Structure)

stan.summ.out <- tmp %>% with(list(fires=Structure[1], casualties=Structure[2], 
                                   batt_fire=Battery[1], batt_cas=Battery[2])) %>%
  sampling(stan.summ, data=., refresh=F)

tmp2 <- stan.summ.out %>% rstan::extract() %>% 
        sapply(function(x) c(mean(x), sd(x), quantile(x, c(0.025, 0.975)))) %>% as.data.frame() %>%
        cbind(value=c("mean", "sd", "lcl", "ucl"), .)
row.names(tmp2) <- NULL
```

This table shows summary information on fires, battery fires, and injuries.

```{r summary-a}
tmp %>% knitr::kable()
```

And this table shows confidence intervals on all the ratios in the table above.
Ratios and confidence intervals for this table are computed in Stan. 
What is significant here is that injuries from battery fires occur at almost 
double the rate of regular structure fires (`r tmp2$r_batt_cas[1] %>% formatC(digits=2, format="f")`, 
Confidence interval: `r tmp2$r_batt_cas[3] %>% formatC(digits=2, format="f")` - 
`r tmp2$r_batt_cas[4] %>% formatC(digits=2, format="f")`).

```{r summary-b}
tmp2[,1:5] %>% knitr::kable()
rm(tmp, tmp2)
```

## Selected (large) departments

Data are from NFIRS. Initially we look at summaries of data for selected 
departments to see what the data looks like.

```{r selected-fig}
selected %>% mutate(fires=total - not_fire) %>% 
             select(state, fdid, fd_name, year, type, fires) %>% 
             pivot_wider(names_from=type, values_from=fires) %>% 
             filter(year > 2018) %>% group_by(fd_name) %>% 
             summarize(Total=sum(Other, na.rm=T), Battery=sum(Battery, na.rm=T)) %>% 
             mutate(Total=Total + Battery, ratio=Total / Battery) %>% 
             arrange(desc(ratio)) %>% 
     knitr::kable(col.names=c("Department", "All Fires", "Battery Fires", "Ratio"))
```

Basically, what is tells me is that NFIRS is terrible when it comes to 
Li ion battery fires. The very wide range of number of battery fires per 
reported fire is not realistically representative of the real world, and can 
only mean that departments vary wildly in the extent to which they identify
battery fires. Worse, the largest departments are typically the worst at it.

## Tesla, by VIN

Specifically looking at car battery fires presents a couple of serious problems.
First, they are very poorly recorded in NFIRS (as seems to be typical of Li ion
battery fires generally). Second, Regular (Pb-acid) battery fires are numerous
in the data. It turns out that the easiest way to identify Li ion battery fires
in automobiles is by identifying the cars by VIN number. At this point, Tesla is
the only maker (and vehicles) that I can specifically identify in the data, so 
this pulls that data.

```{r tesla-est}
figs$figure.01 <- 
      tesla %>% pivot_longer(2:3) %>% ggplot() + aes(x=year, y=value, color=name, group=name) + 
                geom_point() + geom_line() + scale_x_continuous(minor_breaks=2000:2025) + theme_bw()
# This figure is different from what I made before. 
# This will need to be reworked anyway once I get VIN info to work with.
print(figs$figure.01)
```

Only about half of all vehicle fires in NFIRS have VIN numbers associated, so 
expect the number of Li ion battery fires to be at least double these numbers.

# Cars - Action Taken

This looks at action taken for Li ion vehicle fires to see if there is anything 
useful there.

```{r tesla-action, eval=FALSE}
action.codes <- read.csv(paste0(csv_dir, "act_tak_codes.csv"))[,-1]
car.actions[,11:13] %>% as.matrix() %>% as.vector() %>% {.[! is.na(.) & . != '' & . != 'NULL']} %>%
           data.frame(code=.) %>% left_join(action.codes, by=join_by(code==code_value)) %>%
           with(table(code_descr)) %>% sort(decreasing=T)
```

Action taken does not give anything useful. Basically, the only thing it says is
"Put out fire" as the action taken. Not useful.

## List structure battery fires 

What this does for *each* department is fit a Poisson model to the battery fire 
data. Specifically, it fits the following model:

$$B_{dt} \sim Pois\left\{ \lambda_{dt} \right\} $$

where

$$\log\left(\lambda_{dt}\right) = \log\left(F_{dt}\right) + a_d + b_d t$$

That is, the log of the expected proportion of battery fires for each department is 
a linear trend over time.


```{r struc_anal_old, eval=FALSE}
# This needs reworking. I need to NOT filter by the department ratio, since that 
# can cause the very trend I am trying to explain.
#********************************************************
#**  At this point, I think this section is obsolete.  **
#********************************************************

struc.13 <- structures %>% filter(year >= 2013)

t2 <- struc.13 %>% filter(year >= 2018) %>% 
  { aggregate(.[,5:6], list(state=.$state, fdid=.$fdid, fdname=.$fd_name), sum) } %>% 
  mutate(ratio = fires / batteries)
depts <- t2$fdid[t2$ratio < 300]
rm(t2)

outs <- list()
out  <- list()

for(i in depts){
  outs[[i]] <- struc.13 %>% filter(fdid==i) %>% mutate(yr=year - 2013) %>% 
    glm(batteries ~ yr + offset(log(fires)), data=., family=quasipoisson)
  t2 <- summary(outs[[i]])$coefficients %>% as.data.frame() %>% 
    { cbind(dept=i, var=row.names(.), data.frame(., row.names=NULL)) }
  out <- rbind(out, t2)
}

t2 <- out[,1:3] %>% tidyr::pivot_wider(names_from=2, values_from=3)
t2$deptName <- match(t2$dept, struc.13$fdid)
t2$deptName <- struc.13$fd_name[t2$deptName]
t2 <- t2[,c(4,1,2,3)]
t2$ratio <- exp(-t2$`(Intercept)`)

verge <- t2 %>% filter(ratio < 1e5) %>% lm(yr ~ `(Intercept)`, data=.)
ve <- verge %>% vcov() %>% eigen()
tmp <- rnorm(20000) %>% { matrix(., ncol=2) %*% diag(sqrt(ve$values)) %*% t(ve$vectors) }
tmp <- tmp + matrix(rep(coef(verge), 10000), ncol=2, byrow=TRUE)
ver <- c(mean(exp(tmp[,1] / tmp[,2])), sd(exp(tmp[,1] / tmp[,2])))

qq <- c(0.01, 0.025, 0.05, 0.1, 0.5, 0.9, 0.95, 0.975, 0.99)
cur.batt <- data.frame(quantile=qq, 
                  est1=c(mean(-tmp[,1] / tmp[,2]) + qnorm(qq) * sd(-tmp[,1] / tmp[,2])))
cur.batt$rate1 <- exp(-cur.batt$est1)

rm(verge, ve, tmp)
```

```{r struc_anal}
struc.13 <- structures %>% filter(year >= 2013)
depts    <- with(struc.13, paste0(state, ".", fdid)) %>% unique()

outs <- list()
out  <- list()

for(i in depts){
  outs[[i]] <- struc.13 %>% mutate(dept=paste0(state, ".", fdid), yr=year - 2013) %>% filter(dept == i) %>% 
               glm(batteries ~ yr + offset(log(fires)), data=., family=quasipoisson)
  t2 <- summary(outs[[i]])$coefficients %>% as.data.frame() %>% 
    { cbind(dept=i, var=row.names(.), data.frame(., row.names=NULL)) }
  out <- rbind(out, t2)
}

dept_summary <- out[,1:3] %>% tidyr::pivot_wider(names_from=2, values_from=3)
dept_summary$deptName <- match(dept_summary$dept, with(struc.13, paste0(state, ".", fdid)))
dept_summary$deptName <- struc.13$fd_name[dept_summary$deptName]
dept_summary <- dept_summary[,c(4,1,2,3)]
dept_summary$ratio <- exp(-dept_summary$`(Intercept)`)
# Use either the above line OR the following two.
#dept_summary$ratio.2013 <- exp(-dept_summary$`(Intercept)`)
#dept_summary$ratio.2023 <- exp(-dept_summary$`(Intercept)` - 10 * dept_summary$yr)
dept_summary <- struc.13 %>% mutate(dept=paste0(state, ".", fdid)) %>% 
  group_by(dept) %>% summarize(fires=sum(fires), batt=sum(batteries)) %>% merge(dept_summary, .)

verge <- dept_summary %>% filter(batt > 3, `(Intercept)` > -12, `(Intercept)` < -4) %>%
         lm(yr ~ `(Intercept)`, data=.)
ve <- verge %>% vcov() %>% eigen()
tmp <- rnorm(20000) %>% { matrix(., ncol=2) %*% diag(sqrt(ve$values)) %*% t(ve$vectors) }
tmp <- tmp + matrix(rep(coef(verge), 10000), ncol=2, byrow=TRUE)
ver <- c(mean(exp(tmp[,1] / tmp[,2])), sd(exp(tmp[,1] / tmp[,2])))

qq <- c(0.01, 0.025, 0.05, 0.1, 0.5, 0.9, 0.95, 0.975, 0.99)
cur.batt <- data.frame(quantile=qq, 
                  est1=c(mean(-tmp[,1] / tmp[,2]) + qnorm(qq) * sd(-tmp[,1] / tmp[,2])))
cur.batt$rate1 <- exp(-cur.batt$est1)

rm(verge, ve, tmp)
```

```{r figs}
figs$figure.02 <- dept_summary %>% filter(batt > 3, `(Intercept)` > -12, `(Intercept)` < -4) %>% 
   ggplot() + aes(x=`(Intercept)`, y=yr) + geom_point() + geom_smooth(method="lm") + 
   scale_x_continuous()
print(figs$figure.02)
```

Interestingly, the plot of all the slope-intercept values form a linear relationship.
Two models that I can think of can produce this relationship. First, where at some 
time (T), all converge on the same rate. That is, there is a rate *r* and a time *T* 
such that for all departments $d \in \mathcal{D}$, $r = a_d + b_d T$. They converge
at the point where one out of `r as.integer(ver[1])` &pm; `r as.integer(ver[2] * qnorm(0.975))`
fires is a battery fire.

An alternative model that produces this result is the switching model that is developed
in the next two sections. In it there are two (log) linear trends, and each department
switches between the two trends at a random point in the time line. The section below
runs a simulation using this model to verify that it produces a linear trend like 
the one above. The subsequent section attempts to fit such a model 
to the data.

## Testing a model

The idea is that for all departments, the number of reported battery fires follows 
the following model:

$$B_{it} \sim Pois\left(\lambda_{it}\right)$$

where

$$ \ln \lambda_{it} = \left\{ \begin{matrix} a_0 + b_0 t & t < T_i \\ a_1 + b_1 t & t \ge T_i \end{matrix} \right.$$

What I want to do is determine whether this model produces the linear results that 
I see in my data. So I am going to take a take a specific set of $\left\{a_0, a_1, b_0, b_1\right\}$
values and run a bunch of cases, each with different cutoff values ($T_i$), and see 
if it replicates the results I see in the data.

```{r model_data}
model <- list(a0=5 * runif(1) - 15,
              a1=2 * runif(1) - 6,
              b0=0.02 * runif(1) - 0.01,
              b1=0.1 * runif(1),
              F=10000,
              T_max=10,
              n=100)

df <- with(model, data.frame(rep=rep(1:(n * (T_max + 1)), each=T_max),
                             T=rep(0:T_max, each=n * T_max),
                             t=rep(1:T_max, n * (T_max + 1))))

df$llambda <- with(model, ifelse(df$t <= df$T, a0 + df$t * b0, a1 + df$t * b1))
df$fires <- model$F
df$batteries <- rpois(nrow(df), df$fires * exp(df$llambda))

m.results <- data.frame()
for(i in 1:max(df$rep)){
  out.22 <- df %>% filter(rep==i) %>% glm(batteries ~ t + offset(log(fires)), data=., family=quasipoisson) %>% coef()
  m.results <- rbind(m.results, 
                     data.frame(rep=i, intercept=out.22[1], yr=out.22[2]))
}

figs$figure.03 <- m.results %>% filter(intercept > -50, yr > -5) %>% ggplot() + aes(x=intercept, y=yr, color=factor(as.integer(rep/100))) + geom_point()

print(figs$figure.03)
```

The results of the simulation model give results that are largely consistent with what 
we are looking for. With the exception of the smallest battery fire rate, all the
values fall into an approximate straight line, which is consistent with the 
data.

## Estimate Model

The model that we work with is this:

The idea is that reported fires follow a random switching model. Fires follow a 
Poisson distribution where expected number of fires follows one of two log-linear 
time trends. For each department, there is a random time $\tau_i$, where for any 
time before $\tau_i$ expected number of fires follows one trend, and for any time
after $\tau_i$ the expected number follows the other trend.

Formally:

let $\mathcal{D}$ be the set of departments,

$\mathcal{T}$ be the times for which data is collected,

$F_{dt}$ be the number of fires reported for department $d$ at time $t$,

$B_{dt}$ be the number of battery fires reported for department $d$ at time $t$,

and $a_1$, $a_2$, $b_1$, and $b_2$ are parameters.

Finally, for each department $d \in \mathcal{D}$, there is a random switching time 
$\tau_d \in \mathcal{T}$ such that:

$$B_{dt} \sim Pois\left\{ \left\{\begin{matrix} \exp\left(\ln F_{dt} + a_1 + b_1 t\right) & t < \tau_d \\ \exp\left(\ln F_{dt} + a_2 + b_2 t\right) & t \ge \tau_d \end{matrix} \right. \right\}$$

In the code below, the switching time $\tau_d$ is treated as a random variable with
uniform prior.

The version of this model discussed below has each department with its own slope 
and intercept value before switching occurs, but there is a common value for 
after switching occurs. A second version was estimated where both before and 
after trends were common to all departments. Results were not substantially different.

So build a STAN model that will estimate the model.

```{stan model_fit, output.var="stan.model", eval=! exists("stan.model")}
data {
  int                       D;      // Number of departments
  int                       T;      // Number of sample times (years)
  array[D, T] int<lower=0>  batt;   // number of reported battery fires
  array[D, T] int<lower=0>  fires;  // total number of reported fires
  int<lower=0, upper=1>     end;    // 0 - Don't include; 1 - include
  int<lower=0, upper=1>     dept_a; // 0 - Don't include; 1 - include
  int<lower=0, upper=1>     dept_b; // 0 - Don't include; 1 - include
}

transformed data {
  real log_unif = -log(T);
  array[D, T] real  lfires = log(fires);  // log total number of reported fires
  int               a_dim = max(1, dept_a * D);
  int               b_dim = max(1, dept_b * D);
}

parameters {
  vector[a_dim]     a0_par;     // Initial intercept (possibly, by department)
  real              a1;         // Final common intercept
  vector[b_dim]     b0_par;     // Initial slope (possibly, by department)
  real              b1;         // Final common slope
}

transformed parameters {
  matrix[D, T + end] lp;
  {
    vector[D]        a0;
    vector[D]        b0;

    if(a_dim == 1){
      a0 = rep_vector(a0_par[1], D);
    } else {
      a0 = a0_par;
    }
    
    if(b_dim == 1){
      b0 = rep_vector(b0_par[1], D);
    } else {
      b0 = b0_par;
    }

    matrix[D, T + 1] lp_e;
    matrix[D, T + 1] lp_l;
    
    lp_e[:, 1] = rep_vector(0, D);
    lp_l[:, 1] = rep_vector(0, D);
    
    for(d in 1:D){
      for (t in 1:T) {
        lp_e[d, t + 1] = lp_e[d, t] + poisson_log_lpmf(batt[d, t] | lfires[d, t] + a0[d] + b0[d] * t);
        lp_l[d, t + 1] = lp_l[d, t] + poisson_log_lpmf(batt[d, t] | lfires[d, t] + a1 + b1 * t);
      }
      lp[d] = rep_row_vector(log_unif + lp_l[d, T + 1], T + end) + lp_e[d, 1:(T + end)] - lp_l[d, 1:(T + end)];
    }
  }
}

model {
  a0_par ~ normal(0,25);
  b0_par ~ normal(0, 1);
  for(dd in 1:D){
    target += log_sum_exp(lp[dd]);
  }
}

generated quantities {
}
```

```{r stan_dat, eval=! exists("stan.data")}
tmp.f <- struc.13 %>% mutate(dept=paste0(state, ".", fdid)) %>% 
  filter(dept %in% depts) %>% 
  pivot_wider(id_cols=year, names_from=dept, values_from=fires)
tmp.b <- struc.13 %>% mutate(dept=paste0(state, ".", fdid)) %>% 
  filter(dept %in% depts) %>% 
  pivot_wider(id_cols=year, names_from=dept, values_from=batteries)
dd <- tmp.f %>% colSums() %>% is.na() %>% not()
dd[1] <- FALSE

stan.data <- list(D = length(dd[dd]),
                  T = nrow(tmp.f),
                  batt = t(tmp.b[,dd]),
                  fires=t(tmp.f[,dd]),
                  end=0,
                  dept_a=1,
                  dept_b=0)

rm(tmp.f, tmp.b)
```


```{r stan_run, eval=! (exists("stan.out2") & exists("stan.out3"))}
fn <- function(depts.a=1, depts.b=depts.a){
  function(){
    list(a0=rnorm(depts.a, mean=-6, sd=1),
         a1=rnorm(1, mean=-6, sd=1),
         b0=rnorm(depts.b, mean= 0, sd=0.05),
         b1=rnorm(1, mean= 0, sd=0.05))
  }
}
# This provides an initialization function to the model because otherwise
# I end up with garbage results. Doing this is more or less OK because I have 
# a pretty good idea of what values I should end up with.
stan.data$dept_a <- 0   # Common before values
stan.data$dept_b <- 0
stan.out2 <- sampling(stan.model, data=stan.data, init=fn())
stan.data$dept_a <- 1   # Department specific before values
stan.data$dept_b <- 1
stan.out3 <- sampling(stan.model, data=stan.data, init=fn(stan.data$D))
```



```{r posterior}
# This is intended to estimate the posterior probability of a switching time
# for each department.

dpt.temp <- struc.13 %>% mutate(dept=paste0(state, ".", fdid)) %>% 
            filter(dept %in% depts) %>% 
            pivot_wider(id_cols=year, names_from=dept, values_from=fires) %>% 
            colSums() %>% is.na() %>% not()
dpt.temp[1] <- FALSE

dd <- matrix(ncol=11, nrow=0)
lp <- rstan::extract(stan.out3, "lp")
nd <- dim(lp[[1]])[2]
for(i in 1:nd){
  a <- lp[[1]][,i,]
  b <- a %>% apply(2, max)
  a1 <- a - matrix(b, nrow=nrow(a), ncol=ncol(a), byrow=TRUE)
  a2 <- a1 %>% exp() %>% colSums()
  a3 <- log(a2) + b
  a4 <- exp(a3 - max(a3))
  dd <- rbind(dd, a4 / sum(a4))
}

switchdate <- data.frame(dept=1:nd,
                         year=apply(dd, 1, function(x) which(x == max(x))) + 2012)
switchdate$dept <- names(dpt.temp)[dpt.temp]

rm(a, a1, a2, a3, a4, b, i, lp, nd, dd, dpt.temp)
vroom <- rbind(stan.out3 %>% rstan::extract("a1") %>% extract2(1) %>% {c(mean(.), sd(.))},
               stan.out3 %>% rstan::extract("b1") %>% extract2(1) %>% {c(mean(.), sd(.))}) %>%
         as.data.frame() %>% cbind(var=c("intercept", "slope"), .) %>% 
         setNames(c("var", "est", "sd"))
vroom$lcl <- vroom$est + qnorm(0.025) * vroom$sd
vroom$ucl <- vroom$est + qnorm(0.975) * vroom$sd

tuv <- stan.out2 %>% rstan::extract(c("a1", "b1")) %>% sapply(function(x) c(mean(x), sd(x)))
cur.batt$est2 <- tuv[1,1] + 10 * tuv[1,2] + 
                               qnorm(qq) * sqrt(tuv[2,1]^2 + 100 * tuv[2,2]^2)
cur.batt$rate2 <- exp(-cur.batt$est2)

tuv <- stan.out3 %>% rstan::extract(c("a1", "b1")) %>% sapply(function(x) c(mean(x), sd(x)))
cur.batt$est3=tuv[1,1] + 10 * tuv[1,2] + 
                               qnorm(qq) * sqrt(tuv[2,1]^2 + 100 * tuv[2,2]^2)
cur.batt$rate3 <- exp(-cur.batt$est3)

sw <- c(length(switchdate$dept[switchdate$year == 2013]),
        length(switchdate$dept[switchdate$year == 2023]))
```

Initial battery fire rates are mostly indistinguishable from zero for 
all departments. Final value for the 2012 ratio of battery fires to all fires
was one battery fire to `r round(exp(-vroom[1,2]), -1)` fires (95% confidence 
interval: `r round(exp(-vroom[1,5]), -1)`, `r round(exp(-vroom[1,4]), -1)`).
Common final slope was `r round(100 * (exp(vroom[2,2]) - 1), 3)`% per year 
(95% confidence interval: `r round(100 * (exp(vroom[2,4]) - 1), 3)`%, 
`r round(100 * (exp(vroom[2,5]) - 1), 3)`%).

The figure below shows the switchover year for each department in the model.
Some `r sw[1]` departments switched over before the model period began, and `r sw[2]`
switched over in the very last year. The remainder switched over at various times 
throughout the model period.

```{r changeover}
rm(vroom)
figs$figure.04 <- switchdate %>% mutate(year=factor(year)) %>% ggplot() + aes(x=year) +
                 geom_histogram(stat="count") + xlab("Year") + ylab("Count") +
                 theme_bw()
print(figs$figure.04)
```


## Just model the data

Another approach to guessing at the rate of battery fires is to assume that the 
departments in the model report at various random rates (which is fitted by 
a multivariate normal distribution). The actual underlying rate of battery fires
is a quantile of that distribution. One factor is that fires occur randomly. So
if there is randomly an unusually large number of fires early in the sample period
then, by regression to the mean, we would expect a more negative slope, which 
implies that intercept and slope will be negatively correlated. That is to say, 
at least some of the slope in the figure above (figure 1???) is purely due to 
random variation.  This model accounts for that.

```{stan model_var1, output.var="stan.model1", eval=! exists("stan.model1")}
// This estimates the Covariance matrix of the intercept-slope pairs jointly with
// the reported data. Since estimating just the MV Normal for the intercept-slope 
// pair had issues that might fix.
//
// That is, the raw data is estimated using a random-effects model where a 
// slope and intercept value is estimated for each department. Then a MV Normal 
// distribution is fitted to the estimated slope-intercept values.
data {
  int                           N_year;     // Number of years
  int                           N_dept;     // Number of Departments
  array[N_dept, N_year] int     batteries;  // number of reported battery fires
  array[N_dept] vector[N_year]  fires;      // number of reported fires
  vector[N_year]                year;       // year
}

transformed data {
}

parameters {
  array[N_dept]  real  intercept;
  array[N_dept]  real  slope;
  vector[2]            mu;
  cov_matrix[2]        Sigma;
}

transformed parameters {
}

model {
  vector[N_year]   llambda;
  array[N_dept] vector[2]  dept_pars;
  
  for(dept_id in 1:N_dept){
    dept_pars[dept_id, 1] = intercept[dept_id];
    dept_pars[dept_id, 2] = slope[dept_id];
    llambda = log(fires[dept_id]) + rep_vector(intercept[dept_id], N_year) + slope[dept_id] * year;
    batteries[dept_id] ~ poisson_log(llambda);
  }
  dept_pars ~ multi_normal(mu, Sigma);
}

generated quantities {
  vector[2]     svar = eigenvalues_sym(Sigma);
  matrix[2,2]   qq   = eigenvectors_sym(Sigma);
  real          theta1 = acos(qq[1,1]) * 180 / pi();
  real          theta2 = asin(qq[2,1]) * 180 / pi();
  vector[2]     est_90 = qq * diag_matrix(sqrt(svar)) * [0, inv_Phi(0.10 )]' + mu;
  vector[2]     est_95 = qq * diag_matrix(sqrt(svar)) * [0, inv_Phi(0.05 )]' + mu;
  vector[2]     est_97 = qq * diag_matrix(sqrt(svar)) * [0, inv_Phi(0.025)]' + mu;
  vector[2]     est_99 = qq * diag_matrix(sqrt(svar)) * [0, inv_Phi(0.01 )]' + mu;

  vector[N_dept * N_year] log_lik;
  int                     n;
  for (d in 1:N_dept) {
    for (y in 1:N_year){
      n = y + (d - 1) * N_year;
      log_lik[n] = poisson_log_lpmf(batteries[d,y] | log(fires[d,y]) + intercept[d] + slope[d] * year[y]);
    }
  }
}
```

```{r data1}
a <- NULL
for(i in depts) a <- c(a, struc.13 %>% filter(paste0(state, ".", fdid) == i) %>% nrow())
d2 <- depts[a == 11]

struc.13a <- struc.13 %>% mutate(dept = paste0(state, ".", fdid)) %>% 
             filter(dept %in% d2)
st_dat1 <- list(N_year = 11,
                N_dept = length(d2),
                batteries = matrix(struc.13a$batteries, ncol=11, byrow=TRUE),
                fires = matrix(struc.13a$fires, ncol=11, byrow=TRUE),
                year = 0:10)
rm(i, d2, a, struc.13a)
```

```{r model1, eval=! exists("st_out1")}
st_out1 <- sampling(stan.model1, data=st_dat1, iter=4000, warmup=1000)
```


```{r plot-next}
V <- st_out1 %>% rstan::extract("svar") %>% extract2(1) %>% colMeans()  # eigenvalues and
Q <- st_out1 %>% rstan::extract("qq"  ) %>% extract2(1) %>% colMeans()  # eigenvector matrix for the covariance matrix.
mu<- st_out1 %>% rstan::extract("mu"  ) %>% extract2(1) %>% colMeans() # Mean values for the Estimated MV Normal distribution

t2a <- st_out1 %>% rstan::extract(c("intercept", "slope")) %>% sapply(colMeans) %>% as.data.frame() %>% 
       { cbind(dept=1:nrow(.), .) }

# This builds the plot of the data with the error ellipses built from the covariance
# matrix around the data.
ddf <- matrix(c(rep(0, 27), -13:13 / 10), ncol=2) %*% diag(sqrt(V)) %*% t(Q) %>%
       as.data.frame() %>% setNames(c("x", "y"))
ddf$x <- ddf$x + mu[1]
ddf$y <- ddf$y + mu[2]
ddf <- cbind(gp="X", ddf)

dd1 <- matrix(c(-20:20 / 10, rep(0, 41)), ncol=2) %*% diag(sqrt(V)) %*% t(Q) %>%
       as.data.frame() %>% setNames(c("x", "y"))
dd1$x <- dd1$x + mu[1]
dd1$y <- dd1$y + mu[2]
ddf <- rbind(ddf, cbind(gp="Y", dd1))

dd1 <- matrix(qnorm(0.9) * c(cos(pi * -20:20 / 20), sin(pi * -20:20 / 20)), ncol=2) %*%
        diag(sqrt(V)) %*% t(Q) %>% as.data.frame() %>% setNames(c("x", "y"))
dd1$x <- dd1$x + mu[1]
dd1$y <- dd1$y + mu[2]
ddf <- rbind(ddf, cbind(gp="Q90", dd1))

dd1 <- matrix(qnorm(0.95) * c(cos(pi * -20:20 / 20), sin(pi * -20:20 / 20)), ncol=2) %*%
        diag(sqrt(V)) %*% t(Q) %>% as.data.frame() %>% setNames(c("x", "y"))
dd1$x <- dd1$x + mu[1]
dd1$y <- dd1$y + mu[2]
ddf <- rbind(ddf, cbind(gp="Q95", dd1))

figs$figure.05 <- t2a %>% ggplot() + aes(x=intercept, y=slope) + geom_point() + 
                 geom_path(aes(x=x, y=y, group=gp, color=gp), data=ddf)


tuv <- st_out1 %>% rstan::extract(c("mu", "Sigma")) %>% sapply(colMeans)
cur.batt$est4 <- as.vector(c(1, 10) %*% tuv$mu) + 
                          qnorm(qq) * as.vector(c(1, 10) %*% tuv$Sigma %*% c(1, 10))
cur.batt$rate4 <- exp(-cur.batt$est4)
```

```{r figure4}
plot(figs$figure.05)
```

As an alternative, I fit a model with a fixed common growth rate and a random-effects
term for the base rate for each department and compare the two.

```{stan model_var2, output.var="stan.model2", eval=! exists("stan.model2")}
// This estimates the raw data using a random-effects model where an intercept 
// value is estimated for each department. In addition a common slope is estimated
// for the data. Then a Normal distribution is fitted to the estimated intercept 
// values.
data {
  int                           N_year;     // Number of years
  int                           N_dept;     // Number of Departments
  array[N_dept, N_year] int     batteries;  // number of reported battery fires
  array[N_dept] vector[N_year]  fires;      // number of reported fires
  vector[N_year]                year;       // year
}

transformed data {
}

parameters {
  array[N_dept]  real  intercept;
  real                 slope;
  real                 mu;
  real<lower=0>        Sigma;
}

transformed parameters {
}

model {
  vector[N_year]   llambda;

  for(dept_id in 1:N_dept){
    llambda = log(fires[dept_id]) + rep_vector(intercept[dept_id], N_year) + slope * year;
    batteries[dept_id] ~ poisson_log(llambda);
  }
  intercept ~ normal(mu, Sigma);
}

generated quantities {
  vector[N_dept * N_year] log_lik;
  int                     n;
  for (d in 1:N_dept) {
    for (y in 1:N_year){
      n = y + (d - 1) * N_year;
      log_lik[n] = poisson_log_lpmf(batteries[d,y] | log(fires[d,y]) + intercept[d] + slope * year[y]);
    }
  }
}
```




```{r model2}, eval=! exists("st_out2")}
st_out2 <- sampling(stan.model2, data=st_dat1, iter=4000, warmup=1000)
```


The following table compares the fit of the two models. The comparison says that
the first model is a better fit to the data. That is, taking into account the 
possibility that each department is growing at a different rate is a better fit
to the data (even after accounting for the extra 200+ parameters in the model)
than the simplified version of the model. 

```{r compare_models, eval=! exists("loo_loo")}
st_out_ll1 <- extract_log_lik(st_out1, merge_chains = FALSE)
st_out_r_eff1 <- relative_eff(exp(st_out_ll1), cores = 20)
loo1 <- loo(st_out_ll1, r_eff = st_out_r_eff1, cores = 20)

st_out_ll2 <- extract_log_lik(st_out2, merge_chains = FALSE)
st_out_r_eff2 <- relative_eff(exp(st_out_ll2), cores = 20)
loo2 <- loo(st_out_ll2, r_eff = st_out_r_eff2, cores = 20)

loo_loo <- loo_compare(loo1, loo2)
```

```{r looloo}
print(loo_loo)
```

Below are shown the various estimates of the rate at which battery fires occur.
The best estimates range from one battery fire per 60 fires to one battery fire 
per 165 fires. The 95% confidence intervals tend to overlap, with the 
slope-convergence estimate the only one that is substantially different.

While none of these are particularly good estimates and all have very serious 
flaws, the similarity between them gives some hope that in aggregate they provide
some sense of the rate at which battery fires occur.

```{r best_guess}
tuv <- st_out2 %>% rstan::extract(c("mu", "Sigma", "slope")) %>% 
                   sapply(function(x) c(mean(x), sd(x)))
cur.batt$est5 <- tuv[1,1] + 10 * tuv[1,3] + 
                        sqrt(tuv[1,2]^2 + 100 * tuv[2,3]^2) * qnorm(qq)
cur.batt$rate5 <- exp(-cur.batt$est5)

names(cur.batt) <- c("quantile", "est1", "Slope Convergence", "est2", "Switching 1",
                     "est3", "Switching 2", "est4", "MVNorm 1", "est5", "MVNorm2")

figs$figure.06 <- cur.batt %>% t() %>% as.data.frame() %>% setNames(paste0("q.", qq)) %>%
                 as_tibble(rownames="est") %>% filter(! grepl("^[qe]", est)) %>%
                 mutate(est=factor(est, levels=unique(est))) %>% 
                 ggplot() + aes(x=est, y=q.0.5, ymin=q.0.025, ymax=q.0.975) +
                            geom_point() + geom_errorbar(width=0.05) +
                            expand_limits(y=0) + 
                            xlab("Estimate") + ylab("Per Number of Fires") 

figs$figure.06
```

Also below is a table of the same data.

```{r print_table}
cur.batt %>% t() %>% as.data.frame() %>% setNames(paste0("q.", qq)) %>%
     as_tibble(rownames="est") %>% filter(grepl("^[^qe]", est)) %>% 
     magrittr::extract(,c(1,6,9,3)) %>% 
     knitr::kable(col.names=c("Model", "Median", "95% Confidence", "Interval"), digits=0)
```

## Time of Day

```{r tod}
tod.results <- list(
  all= details %>% 
               with(aggregate(data.frame(fires=fdid), 
                              list(year=year - 2015, tod=factor(tod)), length)) %>%
               glm(fires ~ tod + year, data=., family=quasipoisson),
  home=details %>% filter(prop_use %in% c('419', '429')) %>%
               with(aggregate(data.frame(fires=fdid), 
                              list(year=year - 2015, tod=factor(tod)), length)) %>%
               glm(fires ~ tod + year, data=., family=quasipoisson),
  other=details %>% filter(! prop_use %in% c('419', '429')) %>%
               with(aggregate(data.frame(fires=fdid), 
                              list(year=year - 2015, tod=factor(tod)), length)) %>%
               glm(fires ~ tod + year, data=., family=quasipoisson))

tmp.df <- NULL

for(i in names(tod.results)){
  uuu <- tod.results[[i]]
  ndx <- uuu %>% coef() %>% { grep("tod", names(.)) }
  tmp.df <- rbind(tmp.df, 
                  data.frame(group=i,
                       tod=0:24, 
                       val=c(0, uuu %>% coef() %>% { .[ndx] }, 0),
                       se =c(0, uuu %>% vcov() %>% diag() %>% {.[ndx]} %>% sqrt(), 0)))
}

tmp.df$lcl <- with(tmp.df, val + qnorm(0.025) * se)
tmp.df$ucl <- with(tmp.df, val + qnorm(0.975) * se)

rm(uuu, i, ndx)

figs$figure.07 <- tmp.df %>% ggplot() + aes(x=tod, y=val, ymin=lcl, ymax=ucl) +
           geom_point() + geom_line() + geom_ribbon(alpha=0.3) + facet_wrap(~ group) +
           scale_x_continuous(breaks=0:4 * 6, minor_breaks=0:24) + theme_bw()

figs$figure.07
```

Home fires are about 2/3 of the total battery fires in NFIRS. The figure above
shows the effect of time of day on battery fires for all battery fires, home 
battery fires, and battery fires not in homes. All follow a fairly typical 
day-night cycle where battery fires are more common during the day than at night.

Home fires vary from that slightly in that they are more common when people are
active at home. So there is a slight peak in the number of fires between 7 and 9
am, and a much larger peak in the late afternoon or early evening. Non-home 
battery fires have a relatively stable fire rate between the hours of 8am and 
8pm and tend to have a fire rate that is much lower outside of those hours.

## Seasonality

```{r season}
seasonality <- list(month=details %>% filter(prop_use %in% c('419', '429')) %>% 
                      mutate(yr = year - 2015, month=factor(month)) %>% 
                      group_by(yr, month) %>% summarize(fires=n()) %>%
                      glm(fires ~ 0 + yr + month, data=., family=quasipoisson),
                    seas=details %>% filter(prop_use %in% c('419', '429')) %>% 
                      mutate(yr = year - 2015, season=factor(month %/% 3 %>% mod(4))) %>% 
                      group_by(yr, season) %>% summarize(fires=n()) %>%
                      glm(fires ~ 0 + yr + season, data=., family=quasipoisson))

season.df <- summary(seasonality$month)$coefficients %>% as_tibble(rownames="var") %>% as.data.frame() %>%
             magrittr::extract(-1,)
season.df$var <- 1:12
# season.df[1, 2:5] <- c(0, 0, NA, NA)
season.df$var <- factor(season.df$var, labels=month.name)
season.df$lcl <- with(season.df, Estimate + `Std. Error` * qnorm(0.025))
season.df$ucl <- with(season.df, Estimate + `Std. Error` * qnorm(0.975))

tmp <- summary(seasonality$seas)$coefficients %>% as_tibble(rownames="var") %>% as.data.frame() %>%
             magrittr::extract(-1,)
# tmp[1, 2:5] <- c(0, 0, NA, NA)
tmp$var <- 0:3
tmp$lcl <- with(tmp, Estimate + `Std. Error` * qnorm(0.025))
tmp$ucl <- with(tmp, Estimate + `Std. Error` * qnorm(0.975))
names(tmp) <- c("var.a", "Season", "Season.sd", "Season.t", "Season.p", "Season.lcl", "Season.ucl")
uu <- as.integer(season.df$var) %/% 3 %>% mod(4) %>% match(0:3)

season.df$season <- tmp$Season[uu]
season.df$seas.lcl <- tmp$Season.lcl[uu]
season.df$seas.ucl <- tmp$Season.ucl[uu]


figs$figure.08 <- season.df %>% ggplot() + aes(x=var, y=Estimate, ymin=lcl, ymax=ucl, group="Month", color="Month") + 
                 geom_point() + geom_line() + geom_ribbon(alpha=0.3) + 
                 geom_line(aes(y=season, group="Season", color="Season")) +
                 geom_ribbon(aes(ymin=seas.lcl, ymax=seas.ucl, group="Season", color="Season"), alpha=0.3) + 
                 theme_bw()

figs$figure08a <- season.df %>% ggplot() + aes(x=var, y=Estimate, ymin=lcl, ymax=ucl, group="Month", color="Month") + 
                 geom_point() + geom_errorbar(width=0.1) + 
                 geom_step(aes(y=season, group="Season", color="Season"), position=position_nudge(x=-.5)) +
                 geom_ribbon(aes(ymin=seas.lcl, ymax=seas.ucl, group="Season", color="Season"), alpha=0.3) + 
                 theme_bw()

figs$figure.08
```

Looking at season, there is a slight peak in home battery fires over the summer.
Beyond that, it is possible that December battery fires are anomalously high, but 
such results are not statistically significant.

## Severity

```{r sev-base}
sev.base <- list(room=data.summ %>% filter(inc_type == '11', year > 2010) %>% select(-(1:3)) %>% 
                   glm(cbind(room, fires - room)~ 1, data=., family=quasibinomial),
                 struc=data.summ %>% filter(inc_type == '11', year > 2010) %>% select(-(1:3)) %>% 
                   glm(cbind(struc, fires - struc)~ 1, data=., family=quasibinomial),
                 beyond=data.summ %>% filter(inc_type == '11', year > 2010) %>% select(-(1:3)) %>% 
                   glm(cbind(beyond, fires - beyond)~ 1, data=., family=quasibinomial))

sev.base$summary <- sapply(sev.base, 
                           function(x){
                             q <- coef(x)
                             s <- sqrt(vcov(x)[1,1])
                             c(q, s, plogis(q), 
                               plogis(q + qnorm(0.025) * s), 
                               plogis(q + qnorm(0.975) * s))
                             }
                           ) %>% 
       t() %>% as.data.frame() %>% setNames(c("logit(p)", "s", "p", "lcl", "ucl"))
```

```{r sev}
sev.dat <- details %>% filter(! fire_sprd %in% c("", "NULL")) %>% 
           mutate(sev = factor(ifelse(fire_sprd == 3, 4, fire_sprd)))

sev.results <- list(
  all= sev.dat %>% mutate(yr = year - 2015) %>% group_by(yr, sev) %>% 
                   summarize(fires = n()) %>% arrange(desc(sev)) %>% 
                   mutate(batts=cumsum(fires), not=sum(fires) - batts) %>% 
                   filter(sev != '1') %>%
                   glm(cbind(batts, not) ~ sev, data=., family=quasibinomial),
  home=sev.dat %>% filter(prop_use %in% c('419', '429')) %>%
                   mutate(yr = year - 2015) %>% group_by(yr, sev) %>% 
                   summarize(fires = n()) %>% arrange(desc(sev)) %>% 
                   mutate(batts=cumsum(fires), not=sum(fires) - batts) %>% 
                   filter(sev != '1') %>%
                   glm(cbind(batts, not) ~ sev, data=., family=quasibinomial),
  other=sev.dat %>% filter(! prop_use %in% c('419', '429')) %>%
                   mutate(yr = year - 2015) %>% group_by(yr, sev) %>% 
                   summarize(fires = n()) %>% arrange(desc(sev)) %>% 
                   mutate(batts=cumsum(fires), not=sum(fires) - batts) %>% 
                   filter(sev != '1') %>%
                   glm(cbind(batts, not) ~ sev, data=., family=quasibinomial))

sev.df <- NULL

for(i in names(sev.results)){
  uuu <- sev.results[[i]]
  sev.df <- rbind(sev.df, 
                  data.frame(group=i,
                       sev=factor(c("room", "structure", "beyond"),
                                  levels=c("room", "structure", "beyond")), 
                       val=uuu %>% coef(),
                       se =uuu %>% vcov() %>% diag() %>% sqrt()))
}

sev.df$lcl <- with(sev.df, val + qnorm(0.025) * se)
sev.df$ucl <- with(sev.df, val + qnorm(0.975) * se)

rm(uuu, i, sev.dat)

figs$figure.09 <- sev.df %>% mutate(val=plogis(val), lcl=plogis(lcl), ucl=plogis(ucl))%>%
           ggplot() + aes(x=sev, y=val, ymin=lcl, ymax=ucl) +
           geom_point() + geom_path(group=1) + geom_ribbon(group=1, alpha=0.3) +
           facet_wrap(~ group) + theme_bw()

figs$figure.09a <- sev.df %>% filter(group != "all") %>% 
           mutate(val=plogis(val), lcl=plogis(lcl), ucl=plogis(ucl))%>%
           ggplot() + aes(x=sev, y=val, ymin=lcl, ymax=ucl, group=group, color=group) +
           geom_point() + geom_errorbar(width=.05) + 
           xlab("") + ylab("Probability") + theme_bw()


figs$figure.09b <- sev.base$summary[,-3] %>% 
   { cbind(group="baseline", sev=c("room", "structure", "beyond") , .)} %>% 
   mutate(lcl=qlogis(lcl), ucl=qlogis(ucl)) %>% 
   setNames(c("group", "sev", "val", "se", "lcl", "ucl")) %>% 
   rbind(sev.df, .) %>% filter(group != "all") %>% 
   mutate(val=plogis(val), lcl=plogis(lcl), ucl=plogis(ucl))%>%
   ggplot() + aes(x=sev, y=val, ymin=lcl, ymax=ucl, group=group, color=group) +
      geom_point() + geom_errorbar(width=.05) + 
      xlab("") + ylab("Probability") + theme_bw()

figs$figure.09
```

This represents the probability of a battery fire of the size specified or bigger 
for all battery fires, home battery fires, and all battery structure fires that 
are not in homes.

```{r fig6a}
figs$figure.7b
```

This specifically compares those values for home v. not-home fires. Home fires
are more likely to expand to the room of origin (beyond the object of origin)
than other fires. But other fires are more likely to expand beyond the room 
or origin or beyond the structure of origin. It seems likely this reflects
the size of the battery pack that initiated the blaze. Home fires are more 
likely to expand beyond the object of origin but less likely to get bigger 
than that.

## Room in Home

```{r room}
# The following table takes the area of origin field in `details` and 
# consolidates them into a smaller number of (hopefully) more meaningful
# categories.
area_orig <- data.frame(area_orig=sort(unique(details$area_orig)))
area_orig$NewCode <- c("Other", "Hall/Stair", "Exterior", "Hall/Stair", "Entrance/Exit",
                       "Entrance/Exit", "Common Area", "Common Area", "Nonresidential",
                       "Nonresidential", "Common Area", "Nonresidential", "Other",
                       "Bedroom", "Bedroom", "Dining Room", "Kitchen", "Bathroom",
                       "Laundry", "Office", "Bathroom", "Nonresidential", 
                       "Nonresidential", "Nonresidential", "Nonresidential", 
                       "Nonresidential", "Nonresidential", "Other", "Storage", 
                       "Storage", "Closet", "Storage", "Nonresidential", "Other", 
                       "Garage", "Other", "Equipment", "Equipment", "Nonresidential",
                       "Equipment", "Equipment", "Equipment", "Nonresidential", 
                       "Structure", "Structure", "Exterior", "Nonresidential", 
                       "Nonresidential", "Structure", "Exterior", "Nonresidential",
                       "Exterior", "Garage", "Garage", "Garage", "Garage", "Other",
                       "Garage", "Exterior", "Exterior", "Nonresidential", 
                       "Nonresidential", "Exterior", "Exterior", "Nonresidential",
                       "Undetermined")

room.res <- details %>% merge(area_orig, by="area_orig") %>% 
            filter(prop_use %in% c('419', '429')) %>% 
            mutate(yr = year - 2015, rooms=factor(NewCode) %>% relevel("Bedroom")) %>% 
            group_by(yr, rooms) %>% summarize(fires=n()) %>%
            glm(fires ~ yr + rooms, data=., family=quasipoisson)

room.df <- data.frame(room=room.res %>% coef() %>% names(),
                       val=room.res %>% coef(),
                        se=room.res %>% vcov() %>% diag() %>% sqrt())
room.df <- room.df[-2,]
room.df$room[1] <- "Bedroom"
room.df[1,2:3] <- 0
room.df$room <- room.df$room %>% sub("^rooms", "", .)
room.df$lcl <- with(room.df, val + qnorm(0.025) * se)
room.df$ucl <- with(room.df, val + qnorm(0.975) * se)

room.df$room <- factor(room.df$room, 
                       levels=room.df$room[order(room.df$val, decreasing=TRUE)])

figs$figure.0a <- room.df %>% ggplot() + aes(x=room, y=val, ymin=lcl, ymax=ucl) +
            geom_point() + geom_errorbar(width=0.1) + 
            scale_x_discrete(guide = guide_axis(angle = 90)) + xlab("") + ylab("")

figs$figure.0a
```

```{r play-with-rankings}
N <- 1000
mu <- coef(room.res)[-(1:2)]
Sigma <- vcov(room.res)[-(1:2), -(1:2)] %>% eigen()
nv <- length(mu)
nm <- c("Bathroom", names(mu) %>% sub("^rooms", "", .))

vals <- rnorm(nv * N) %>% matrix(ncol=N)
vals <- mu + Sigma$vectors %*% diag(sqrt(Sigma$values)) %*% vals
vals <- rbind(0, vals)

ranks <- as.integer(NA) %>% matrix(ncol=N, nrow=nv + 1)
for(i in 1:N){ ranks[,i] <- rank(vals[,i], ties.method="first") }
ranks <- ranks %>% t() %>% as.data.frame() %>% setNames(nm)

rank.list <- as.integer(0) %>% matrix(ncol=nv + 1, nrow=nv + 1)
colnames(rank.list) <- nm
row.names(rank.list) <- 1:(nv + 1)

for(i in names(ranks)){
  uuu <- ranks[[i]] %>% table() %>% { c(as.integer(names(.)), as.integer(.)) } %>% matrix(ncol=2)
  rank.list[uuu[,1], i] <- uuu[,2]
}

# I am probably not going to print this out, but I will keep it in case
# it ever becomes useful. The discussion below relies on this, though,
# for some of its conclusions.

rm(uuu, vals, ranks, i, nv, nm, N, mu, Sigma)
```

The garage is the number one area of origin for battery fires, followed by the 
bedroom and common areas. The kitchen is fourth, with other areas following after
those.

```{r cleanup-02}
# Save the relevant file, as below. Then delete everything if you are working in bookdown.
# That means that a `bookdown` flag needs to be set somehow. And then unset at the end.
save(list=intersect(keep2, ls()), file=paste0(intmdt_dir, "NFIRS.RData"))
rm(list=setdiff(ls(), keep))
```