---
title: "Batteries"
subtitle: "CPSC Data"
output: html_document
date: "2024-12-13"
---

# CPSC

```{r setup-03, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
library(magrittr)
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(foreach)
library(textmineR)
library(maxLik)
library(reticulate)
library(ggtern)

keep <- c(ls(), "figs", "fire.reports", "cpsc.ai")

if(! exists("csv_dir")){
  csv_dir <- "csv/"
}
if(! exists("intmdt_dir")){
  intmdt_dir <- "rdata/"
}

# Either load the completed data (currently in .RData, but henceforth in NFIRS.RData) 
# or the relevant compiled data in NFIRS.Data.RData.
ff <- list.files(intmdt_dir) %>% grep("RData", ., ignore.case=T, value=T)
if("cpsc.rdata" %in% tolower(ff)){
  fg <- "cpsc.rdata" %>% match(tolower(ff)) %>% { ff[.] }
} else if("cpsc.data.rdata" %in% tolower(ff)){
  fg <- "cpsc.data.rdata" %>% match(tolower(ff)) %>% { ff[.] }
} else {
  stop("The data files needed for the NFIRS analysis are not present!")
}
load(paste0(intmdt_dir, fg))
rm(ff, fg)

keep2 <- c("incidents", "incident.txt", "incident.dtm1", "cluster_summary", "opt.k",
           "incident.lda", "incident.topics", "fire.summ", "fire.summ.lm", 
           "incidents.fire", "ai.results", "ai.summ")




source("load.files.R")
if(load.data(final=c("cpsc.RData"), 
             data =c("cpsc.data.RData"), 
             dir=intmdt_dir) == 1) stop("The data files needed for the Ca Fires analysis are not present!")

rm(load.data)

keep2 <- list(cpsc=c("incidents", "incident.txt", "incident.dtm1", "cluster_summary", "opt.k",
                     "incident.lda", "incident.topics", "fire.summ", "fire.summ.lm", 
                     "incidents.fire", "ai.results", "ai.summ")
)
if(! exists("figs")) figs <- list()
```

```{r v-time}
figs$figure.10 <- lubridate::mdy(incidents$Report.Date) %>% data.frame(date=lubridate::year(.)) %>%
                  filter(date >= 2010) %>% ggplot() + aes(x=date) + geom_bar()
print(figs$figure.10)
```

There is no obvious trend over time for this data set.

This uses an AI tool to recognize which incidents have fires, and whether there
was a Li-Ion battery involved. Details of the approach are described in the 
nearby table.

```{r ai1, eval=! exists("ai.results")}
fire_reports_ai <- incidents$Incident.Description
```


```{python ai-setup}
import httpx
import os
import re

model = "Llama-4-Maverick-17B-128E-Instruct-FP8"

sys_text = """The text is a description of an incident involving a defective 
or potentially defective product. The text was submitted by a consumer 
to the Consumer Product Safety Commission as part of their consumer 
product monitoring program.

For the text please briefly answer the following question.

QQQ
"""

question = {
  "lithium": "Does this incident involve a lithium ion battery?",
  "fire": "Did a fire occur?"
}

replicates = 10

```

```{python ai-connect, eval=! exists("ai.results")}
import openai
# import numpy as np

test_client = openai.OpenAI(
  base_url=os.environ.get("RCHAT_URL"),
  api_key=os.environ.get("RCHAT_API_KEY"),
  http_client=httpx.Client(verify=False)
)
```

```{python ai-run, eval=! exists("ai.results")}
report = r.fire_reports_ai

output = dict()

for u, v in question.items():
  qq = re.sub("QQQ", v, sys_text)
  outey = []  
  for j in range(reps): 
    outex = []
    for i in range(len(report)):
      try:
        response = test_client.chat.completions.create(
          model=model,
      #    max_tokens=128,
      #    temperature=0.7,
      #    top_p=0.95,
      #    stream=False,
          messages=[
            {"role": "system", "content": qq},
            {"role": "user", "content": report[i]}
          ]
        )
        outex.append(reponse.choices[0].message.content)
      except Exception as e: 
      # deal with errors, but for now, do nothing
#        raise e
        outex.append("Error")
    #  print(response.choices[0].message.content)
    outey.append(outex)
  output[u] = outey
```

```{r method-table}
c("AI Model", py$model,
  "System Prompt", py$sys_text,
  "Lithium Question", py$question[[1]],
  "Fire Question", py$question[[2]],
  "Replicates", py$replicates
  ) %>% matrix(ncol=2, byrow=T) %>% as.data.frame() %>% setNames(c("Item", "Value")) %>% 
  knitr::kable()
```

Because Large Language Models are inherently stochastic, each question was repeated
ten times for each incident. The nearby table summarizes the results. The 'Yes' 
(resp, 'No') column represents the number of incidents in which all replicates 
agreed that the answer to the question was 'Yes' (resp. 'No'). The 'Unknown' column 
represents incidents where the replicates did not agree on an answer or where 
some replicates did not give a 'yes' or 'no' answer. An example of the latter 
is an incident where the AI give the following response:

> To answer your question: The text does not explicitly state that the incident 
involves a lithium-ion battery. It only mentions "the battery." Therefore, based 
on the information provided, we cannot confirm whether the incident involves a 
lithium-ion battery or not.

For the 'Combined' row, 'No' represents incidents where either Lithium or Fire
is no, and 'Yes' represents incidents where both Lithium and Fire are 'Yes'. 
The remaining rows include either the Fire unknowns or the Lithium unknowns in
the Yes column.

```{r processing, eval=! exists("ai.results")}
ai.results <- py$output %>% lapply(function(x) do.call("cbind", x))
# This bit gets the 'yes/no' responses out of the text discussion.
ai.summ <- ai.res0 %>% lapply(function(x) str_extract(x, "(^|:[[:space:]*]*)([Yy]es|[Nn]o)") %>%
                                          str_replace_all("[[:space:]*:]", "") %>% 
                                          matrix(ncol=10))
```

```{r ai-summary-table}
ai.summ1 <- ai.summ %>% lapply(function(y) apply(y, 1, function(x) x %>% replace_na("Unknown") %>% 
                                       { c(length(.[. == "Yes"]), length(.[. == "No"]))} ) %>% t() %>% 
                                       as.data.frame() %>% setNames(c("Yes", "No")) %>% 
                                       mutate(result=ifelse(Yes == 10, "Yes", 
                                                            ifelse(No == 10, "No", "Unknown"))))
ai.summary.table <- rbind(
  ai.summ1$lithium$result %>% table(),
  ai.summ1$fire$result %>% table(),
  with(ai.summ1, ifelse(lithium$result == "No" | fire$result == "No", "No",
                 ifelse(lithium$result == "Unknown" | fire$result == "Unknown", "Unknown", "Yes"))) 
       %>% table(),
  with(ai.summ1, ifelse(lithium$result == "No" | fire$result == "No", "No",
                 ifelse(fire$result == "Unknown", "Unknown", "Yes"))) 
       %>% table(),
  with(ai.summ1, ifelse(lithium$result == "No" | fire$result == "No", "No",
                 ifelse(lithium$result == "Unknown", "Unknown", "Yes"))) 
       %>% table()
) %>% as.data.frame() %>% cbind(data=c("Lithium", "Fire", "Combined", "Combined - Include Li Unk", "Combined - Include Fire Unk"), .)

knitr::kable(ai.summary.table)
```

This chart is a ternary plot of the responses where answers were ambiguous.

```{r ai-unknowns-plot}
figs$ternary <- rbind(cbind(Question="Lithium", ai.summ1$lithium),
      cbind(Question="Fire",    ai.summ1$fire)) %>% 
   filter(Yes != 10 & No != 10) %>% 
  mutate(Unknown=10 - Yes - No, 
         Unknown=0.5 + Unknown * 0.9,
         Yes=0.5 + Yes * 0.9,
         No=0.5 + No * 0.9) %>% 
   ggtern(aes(x=Yes, y=No, z=Unknown)) + geom_density_tern() + facet_wrap(~ Question)

print(figs$ternary)
```


```{r create-fires, eval=! exists("cpsc.ai")}
cpsc.ai <- cbind(incidents[,c("Report.Date", "City", "State")],
                 st_abb=as.character(NA),
                 incidents[,c("ZIP", "Location", "X.Primary..Victim.Severity")], 
                 Lithium=as.logical(NA),
                 Lithium.all=as.logical(NA),
                 Fire=as.logical(NA),
                 Fire.all=as.logical(NA),
                 Description=incidents[,"Incident.Description"]) %>% 
           rename(Severity = X.Primary..Victim.Severity)%>% 
  mutate(Report.Date=lubridate::mdy(Report.Date),
         Severity=ifelse(grepl("^Injury", Severity), "Injury", Severity),
         Severity=ifelse(grepl("^Incident", Severity), "Incident", Severity),
         Severity=ifelse(grepl("Unspecified", Severity), "Unspecified", Severity),
         Lithium=ai.summ1$lithium$Yes == 10,
         Lithium.all=ai.summ1$lithium$No != 10,
         Fire=ai.summ1$fire$Yes == 10,
         Fire.all=ai.summ1$fire$No != 10)
cpsc.ai$st_abb <- states$abb[match(tolower(cpsc.ai$State), tolower(states$name))]
# These have been identifed by hand.
cpsc.ai$st_abb[with(cpsc.ai, Lithium & Fire & is.na(st_abb) & ! is.na(ZIP))] <- 
      c("OH", NA, "TX", "CA", NA, "PA", NA, NA, "CA", NA, NA)
```

```{r ai-trend-chart0}
s1 <- cpsc.ai %>% mutate(Year=lubridate::year(Report.Date)) %>% 
    filter(Lithium, Fire, Year < 2025) %>% 
    group_by(Year) %>% summarize(Fires=n()) %>% lm(log(Fires) ~ I(Year - 2010), data=.)
s2 <- summary(s1)$coefficients[2,1:2] 
s2 <- c(s2, s2[1] + qnorm(c(0.025, 0.975)) * s2[2]) * 100
```

Below shows the trend of reported Li ion battery fires over time. The number of 
reported fires is increasing at the rate of about `r formatC(s2[1], digits=2)`&nbsp;% 
(95&nbsp;% Confidence Interval: 
`r formatC(s2[3], format="f", digits=1)` - `r formatC(s2[4], format="f", digits=1)`
per year.

```{r ai-trend-chart1}
figs$figure.12a <- cpsc.ai %>% mutate(Year=lubridate::year(Report.Date)) %>% 
    filter(Lithium, Fire, Year < 2025) %>% 
    group_by(Year) %>% summarize(Fires=n()) %>% 
    ggplot() + aes(x=Year, y=Fires) + geom_point() + geom_smooth(method='lm') + 
        expand_limits(x=2010, y=10) + scale_y_log10(breaks=1:9 * 10, minor_breaks=NULL) +
        scale_x_continuous(breaks=2010 + 0:3 * 5, minor_breaks=2010:2025) +
        theme_bw()

print(figs$figure.12a)
```

```{r cleanup-03}
# Save the relevant file, as below. Then delete everything if you are working in bookdown.
# That means that a `bookdown` flag needs to be set somehow. And then unset at the end.
for(i in names(keep2)){
  save(list=intersect(keep2[[i]], ls()), file=paste0(intmdt_dir, i, ".RData"))
}
rm(list=setdiff(ls(), keep))
```