---
title: "Batteries"
subtitle: "CPSC Data"
output: html_document
date: "2024-12-13"
---

# CPSC

```{r setup-03, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
library(magrittr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(foreach)
library(textmineR)
library(maxLik)

keep <- c(ls(), "figs", "fire.reports")

if(! exists("csv_dir")){
  csv_dir <- "csv/"
}
if(! exists("intmdt_dir")){
  intmdt_dir <- "rdata/"
}

# Either load the completed data (currently in .RData, but henceforth in NFIRS.RData) 
# or the relevant compiled data in NFIRS.Data.RData.
ff <- list.files(intmdt_dir) %>% grep("RData", ., ignore.case=T, value=T)
if("cpsc.rdata" %in% tolower(ff)){
  fg <- "cpsc.rdata" %>% match(tolower(ff)) %>% { ff[.] }
} else if("cpsc.data.rdata" %in% tolower(ff)){
  fg <- "cpsc.data.rdata" %>% match(tolower(ff)) %>% { ff[.] }
} else {
  stop("The data files needed for the NFIRS analysis are not present!")
}
load(paste0(intmdt_dir, fg))
rm(ff, fg)

keep2 <- c("incidents", "incident.txt", "incident.dtm1", "cluster_summary", "opt.k",
           "incident.lda", "incident.topics", "fire.summ", "fire.summ.lm", 
           "incidents.fire")
```

```{r v-time}
figs$figure.10 <- lubridate::mdy(incidents$Report.Date) %>% data.frame(date=lubridate::year(.)) %>%
                  filter(date >= 2010) %>% ggplot() + aes(x=date) + geom_bar()
print(figs$figure.10)
```

There is no obvious trend over time for this data set.

```{r cpsc, eval=! exists("cluster_summary")}
incident.dtm <- CreateDtm(incident.txt,
                          ngram_window=c(1,2),
                          stopword_vec=c(stopwords::stopwords("en"), 
                                         stopwords::stopwords(source = "smart"),
                                         "redacted"),
                          stem_lemma_function=SnowballC::wordStem)

tf_mat <- TermDocFreq(incident.dtm)

# What follows is a bastardized version of:
# > incident.dtm1 <- incident.dtm[,tf_mat$doc_freq > 4]
# The reason I do this is because the above command is HIGHLY inefficient
# taking forever and using huge amounts of memory. This, as annoying as it
# is, works better, even counting the time required to build the routine.
#
# The idea is get rid of any entries that occur in fewer than 5 documents.
# They would contribute nothing to the analysis, and just consume computational
# resources.
i <- NULL
x <- NULL
m <- NULL
p <- 0
ll <- apply(incident.dtm, 2, function(x) sum(x > 0))
for(j in 1:ncol(incident.dtm)){
  if(ll[j] > 4){
    i <- c(i, incident.dtm@i[(incident.dtm@p[j] + 1):incident.dtm@p[j + 1]])
    x <- c(x, incident.dtm@x[(incident.dtm@p[j] + 1):incident.dtm@p[j + 1]])
    m <- c(m, incident.dtm@Dimnames[[2]][j])
    p <- c(p, length(x))
  }}

incident.dtm1 <- incident.dtm
incident.dtm1@i <- i
incident.dtm1@p <- as.integer(p)
incident.dtm1@x <- x
incident.dtm1@Dim[2] <- length(m)
incident.dtm1@Dimnames[[2]] <- m

####### End annoying section ##########

tf_mat <- TermDocFreq(incident.dtm1)

tfidf <- incident.dtm1[,tf_mat$term] %>% t() %>% multiply_by(tf_mat$idf) %>% t()

csim <- (tfidf / sqrt(rowSums(tfidf * tfidf))) %>% { . %*% t(.)}
cdist <- 1 - csim

hc <- hclust(as.dist(cdist), "ward.D")

N_max <- 30
clustering <- list()
silhouette <- rep(NA, N_max)
for(i in 1:N_max){
  clustering[[paste0("n = ", 5 * i)]] <- cutree(hc, 5 * i)
  M <- matrix(rep(0, 3800 * 5 * i), ncol=5 * i)
  M[cbind(1:3800, clustering[[i]])] <- 1
  
  tmp <- cdist %*% M %*% diag(1 / colSums(M))
  t3 <- rep(NA, nrow(M))
  for(j in 1:(5 * i)){
    ndx <- as.logical(M[,j])
    sc <- sum(ndx) / (sum(ndx) - 1)
    t2 <- apply(tmp[ndx,-j], 1, min)
    t3[ndx] <- (t2 - sc * tmp[ndx,j]) / pmax(t2, sc * tmp[ndx,j])
  }
  silhouette[i] <- mean(t3)
}
rm(i, j, N_max, ndx, sc, t2, t3, M, tmp, csim)

#####################

p_words <- colSums(incident.dtm1) / sum(incident.dtm1)

cluster_words <- lapply(unique(clustering[[4]]), function(x){
  rows <- incident.dtm1[ clustering[[4]] == x , ]
  
  # for memory's sake, drop all words that don't appear in the cluster
  rows <- rows[ , colSums(rows) > 0 ]
  
  colSums(rows) / sum(rows) - p_words[ colnames(rows) ]
})

# create a summary table of the top 5 words defining each cluster
cluster_summary <- data.frame(cluster = unique(clustering[[4]]),
                              size = as.numeric(table(clustering[[4]])),
                              top_words = sapply(cluster_words, function(d){
                                paste(
                                  names(d)[ order(d, decreasing = TRUE) ][ 1:5 ], 
                                  collapse = ", ")
                              }),
                              stringsAsFactors = FALSE)
```

First we use some basic text processing tools to cluster the reports. This
was grouped into 20 clusters, for which basic details are shown in the table below. Interestingly, most of the clusters are associated with specific brands 
of devices. Two are related to smoke alarms. One cluster is basically a catch-all
that captures all the rest of the reports.

```{r cs}
cluster_summary %>% knitr::kable()
```


Using Latent Dirichlet Allocation (LDA) topics are identified in the reports.



```{r find, eval=! exists("opt.k")}
# opt.k <- data.frame(k=1:10 * 10, coherence=NA)
#for(i in 1:nrow(opt.k)){
#  system.time(
#    tmp <- FitLdaModel(incident.dtm1, 
#                       k=opt.k$k[i],
#                       iterations=500, 
#                       burnin=400, 
#                       optimize_alpha=T, 
#                       calc_coherence=T))
#  opt.k$coherence[i] <- mean(tmp$coherence)
#}
cl <- parallel::makeCluster(10)
doParallel::registerDoParallel(cl)
opt.k <- foreach(k==1:10 * 10, .combine="rbind") %dopar% {
  a <- system.time({
    tmp <- FitLdaModel(incident.dtm1, 
                       k=opt.k$k[i],
                       iterations=500, 
                       burnin=400, 
                       optimize_alpha=T, 
                       calc_coherence=T)
    })
  c(k=k, coherence=mean(tmp$coherence), dt=a[3])
}
parallel::stopCluster(cl)
opt.k <- as.data.frame(opt.k)
rm(cl)
```

```{r opt.k}
figs$figure.11 <- opt.k %>% ggplot() + aes(x=k, y=coherence) + geom_point() + geom_line()
print(figs$figure.11)
```

The above chart suggests that about 50 topics is best. Using LDA, the 
following table lists the 50 topics.

```{r lda, eval=! exists("incident.lda")}
incident.lda <- FitLdaModel(incident.dtm1, 
                            k=50, 
                            iterations=1000, 
                            burnin=500, 
                            optimize_alpha=T, 
                            calc_likelihood=T, 
                            calc_coherence=T, 
                            calc_r2=T)
```

```{r topics, eval=! exists("incident.topics")}
incident.topics <- data.frame(topic=1:50, 
                              coherence=incident.lda$coherence,
                              docs=incident.lda$theta %>% apply(2, function(x)
                                                                   length(x[x > 0.1])),
                              terms=GetTopTerms(incident.lda$phi, 5) %>% 
                                    apply(2, function(x) paste0(x, collapse=", ")) %>%
                                    gsub("i,", "y,", .) %>% gsub("i[_]", "y_", .) %>%
                                    gsub("i$", "y", .))
```

```{r print_topics}
incident.topics %>% knitr::kable()
```

```{r fire_trend}
fire.reports <- incident.lda$theta[,c(10,19,25,26,45)] %>%
    magrittr::is_greater_than(0.1) %>% 
    apply(1, function(x) c(10,19,25,26,45)[x] %>% paste0(collapse=", "))
fire.reports[nchar(fire.reports) == 0] <- NA

fire.summ <- incidents$Report.Date[! is.na(fire.reports)] %>% lubridate::mdy() %>%
      data.frame(year=lubridate::year(.)) %>% filter(year > 2010, year < 2025) %>%
      group_by(year) %>% summarize(incidents=n()) 

fire.summ.lm <- lm(log(incidents) ~ I(year - 2010), fire.summ)

val <- summary(fire.summ.lm)$coefficients[2,1:2]
val <- c(est=exp(val[1]) - 1,
         se=(exp(val[1] + val[2] * qnorm(0.975)) - 
                      exp(val[1] + val[2] * qnorm(0.025))) / 2,
         lcl=exp(val[1] + val[2] * qnorm(0.025)) - 1,
         ucl=exp(val[1] + val[2] * qnorm(0.975)) - 1)

fire.summ <- cbind(fire.summ, predict(fire.summ.lm, interval="conf"))
fire.summ$fit <- exp(fire.summ$fit)
fire.summ$lwr <- exp(fire.summ$lwr)
fire.summ$upr <- exp(fire.summ$upr)

figs$figure.12 <- fire.summ %>% ggplot() + aes(x=year, y=incidents, ymin=lwr, ymax=upr) + 
    geom_point() + geom_line() + expand_limits(y=0, x=2010) + 
    geom_point(aes(y=fit), color="blue") + geom_line(aes(y=fit), color="blue") + 
    geom_ribbon(alpha=0.3, color="blue") +
    scale_x_continuous(breaks=0:3 * 5 + 2010, minor_breaks=2010:2025) +
    theme_bw()
print(figs$figure.12)
```

The above chart looks at the trend over time for the number of reports, looking 
specifically at those reports associated with topics related to fire. Specifically
those are the topics numbered 10, 19, 25, 26, and 45. 

There is a marginally significant growth rate of `r formatC(val[1] * 100, digits=2)`
&pm; `r formatC(val[2] * 100, digits=2)`&nbsp;% per year in the number of 
reports in the CPSC database associated with these fire-related topics.

```{r injuries-cpsc, eval=FALSE}
incidents.fire <- incidents[! is.na(fire.reports), 
                            c("Report.Date", "City", "State",
                              "X.Primary..Victim.Severity", 
                              "Incident.Description")] %>% 
  setNames(c("Date", "City", "State", "Old.Sev", "Description")) %>% 
  mutate(Severity="No Incident",
         Severity=ifelse(grepl("Death", Old.Sev), "Death", Severity),
         Severity=ifelse(grepl("^Injury", Old.Sev), "Injury", Severity),
         Severity=ifelse(grepl("^Incident", Old.Sev), "Incident", Severity),
         Severity=ifelse(grepl("Unspecified", Old.Sev), "Unspecified", Severity))
```


```{r cleanup-03}
# Save the relevant file, as below. Then delete everything if you are working in bookdown.
# That means that a `bookdown` flag needs to be set somehow. And then unset at the end.
save(list=intersect(keep2, ls()), file=paste0(intmdt_dir, "cpsc.RData"))
rm(list=setdiff(ls(), keep))
```