---
title: "Batteries"
subtitle: "Capture-Recapture Estimates"
output: html_document
date: "2024-12-13"
---

# Combined NFIRS and CPSC

```{r setup-04, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
#
# For now, keep the old methods in, but soon, delete them and leave
# just the ai-screened set.
#
library(magrittr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(rstan)

keep <- c(ls(), "figs")

if(! exists("csv_dir")){
  csv_dir <- "csv/"
}
if(! exists("intmdt_dir")){
  intmdt_dir <- "rdata/"
}

# Either load the completed data from recapture.rdata
# or the relevant compiled data in recapture.data.rdata.
ff <- list.files(intmdt_dir) %>% grep("RData", ., ignore.case=T, value=T)
if("recapture.rdata" %in% tolower(ff)){
  fg <- "recapture.rdata" %>% match(tolower(ff)) %>% { ff[.] }
} else if("recapture.data.rdata" %in% tolower(ff)){
  fg <- "recapture.data.rdata" %>% match(tolower(ff)) %>% { ff[.] }
} else {
  stop("The data files needed for the Recapture analysis are not present!")
}
load(paste0(intmdt_dir, fg))
rm(ff, fg)

keep2 <- c("incidents", "details", "cpsc.wk", "fire.reports", "overlap", "yearly",
           "cpsc.wk1", "overlap1", "stan.con.model", "stan.con.mdl1", "stan.con.fit",
           "stan.con.fit1", "yearly.stan", "incident.lda", "cpsc.ai", "overlap.ai")

ff <- c("incident.lda", "fire.reports", "cpsc.ai") %in% ls()
if(! all(ff)){
  e <- new.env()
  load(paste0(intmdt_dir, "cpsc.RData"), e)
  incident.lda <- e$incident.lda
  fire.reports <- e$fire.reports
  cpsc.ai      <- e$cpsc.ai
  rm(e)
}
```

## Connect to NFIRS

```{r connect-prep, eval=FALSE}
cpsc.wk <- incidents[! is.na(fire.reports), c("Report.Date", "City", "State", "ZIP",
                                              "Location", "X.Primary..Victim.Severity",
                                              "Incident.Description")] %>% 
  rename(Severity = X.Primary..Victim.Severity) %>% 
  mutate(st_abb=as.character(NA),
         Severity=ifelse(grepl("^Injury", Severity), "Injury", Severity),
         Severity=ifelse(grepl("^Incident", Severity), "Incident", Severity),
         Severity=ifelse(grepl("Unspecified", Severity), "Unspecified", Severity))
cpsc.wk$st_abb <- states$abb[match(tolower(cpsc.wk$State), tolower(states$Name))]
cpsc.wk <- cpsc.wk %>% mutate(st_abb=ifelse(is.na(st_abb), 
                                    ifelse(State %in% states$abb, 
                                           State,
                                           ifelse(State == "District of Columbia", 
                                                  "DC", NA)), 
                                    st_abb),
                              Report.Date=lubridate::mdy(Report.Date))

# This set of corrections is the product of painstakingly going through the 
# cases where there is no valid state, but a city name exists. Then looking up
# the ZIP codes and checking.
t2 <- cpsc.wk %>% filter(is.na(st_abb), ! is.na(City)) %>% row.names()
cpsc.wk$st_abb[row.names(cpsc.wk) %in% t2] <- c(NA, NA, "NY", 'ME', "NM", NA, "MI",
                                                "CA", NA, NA, "CA", NA, NA, "WI", NA,
                                                "FL", NA, NA, "TX", NA, "UT", "KS",
                                                "CA", "CA", "SD", "PA", NA, "NY", NA,
                                                NA, "NC", NA, "CA", "PA", NA, NA, NA,
                                                "WV", NA, "NC")
rm(t2)

cpsc.wk <- cpsc.wk %>% filter(! is.na(st_abb))
```


```{r connecting, eval=! exists("overlap")}

tmp <- match(with(cpsc.wk, paste0(City, ", ", st_abb, ": ",
                                  Report.Date %>% lubridate::year()) %>% tolower()), 
             with(details, paste0(city, ", ", state_id, ": ", year) %>% tolower()))

overlap <- cbind(cpsc.wk %>% {.[(1:nrow(cpsc.wk))[! is.na(tmp)],
                                c("City", "st_abb", "ZIP", "Report.Date", 
                                  "Location", "Severity")]},
                 details[tmp[! is.na(tmp)],c(12:14,5,7)])
```


```{r alternative}
# I didn't trust the 0.1 cutoff for the LDA topic analysis. So this 
# goes through several different approaches to including topics.
# The question is whether this produces different results than 
# the approach described above.
#
# The answer is that it does not, but this documents that process
# so it can be double-checked as necessary.

# This uses four different criteria for including a topic.
#  a.25: ANY of the topics tops a 25% rating
#  s.25: The SUM of the topics tops a 25% rating
#  s.50: The SUM of the topics tops a 50% rating
#  m   : Any of the topics is the top topic associated with that report
cpsc.wk1 <- list(a.25=apply(incident.lda$theta[,c(10,19,25,26,45)], 1, 
                            function(x) any(x > 0.25)) %>% 
                      {.[! is.na(fire.reports)]} %>% {cpsc.wk[.,]},
                 s.25=apply(incident.lda$theta[,c(10,19,25,26,45)], 1, sum) %>%
                      is_greater_than(0.25) %>% {.[! is.na(fire.reports)]} %>%
                      {cpsc.wk[.,]},
                 s.50=apply(incident.lda$theta[,c(10,19,25,26,45)], 1, sum) %>%
                      is_greater_than(0.5) %>% {.[! is.na(fire.reports)]} %>%
                      {cpsc.wk[.,]},
                 m=apply(incident.lda$theta, 1, function(x) match(max(x), x)) %>%
                      is_in(c(10,19,25,26,45)) %>% {.[! is.na(fire.reports)]} %>%
                      {cpsc.wk[.,]})

tmp <- lapply(cpsc.wk1, function(x) match(with(x, paste0(City, ", ", st_abb, ": ",
                  Report.Date %>% lubridate::year()) %>%
                  tolower()),with(details, paste0(city, ", ", state_id, ": ", year) %>%
                                    tolower())))

overlap1 <- list()
for(i in names(tmp)){
   overlap1[[i]] <- cbind(cpsc.wk1[[i]] %>% {.[(1:nrow(cpsc.wk1[[i]]))[! is.na(tmp[[i]])],
                                         c("City", "st_abb", "ZIP", "Report.Date", 
                                           "Location", "Severity")]},
                          details[tmp[[i]][! is.na(tmp[[i]])],c(12:14,5,7)])
}

tmp <- list()
for(i in names(cpsc.wk1)){
 tmp[[i]] <- data.frame(year=2011:2023,
                      cpsc=cpsc.wk1[[i]] %>% mutate(year=lubridate::year(Report.Date)) %>%
                        filter(year < 2024) %>% with(table(year)) %>% as.vector(),
                      nfirs=yearly$nfirs[1:13],
                      overlap=overlap1[[i]]$Report.Date %>% lubridate::year() %>%
                        factor(levels=2011:2023) %>% table() %>% as.vector()) %>% 
              rbind(., colSums(.)) %>% cbind(test=i, .)
}

tmp <- tmp %>% do.call("rbind", .) %>% mutate(N=cpsc * nfirs / overlap)
# This plots all the different versions. The baseline (which is what is reported)
# falls near the middle. This plot should probably not be in the report.
figs$figure.20 <- tmp %>% filter(year < 2100, N < Inf) %>% ggplot() + 
  aes(x=year, y=N, color=test, group=test) + geom_point() + geom_line() +
  geom_point(aes(x=Year, y=N, color="base", group="base"), data=yearly[1:13,]) +
  geom_line(aes(x=Year, y=N, color="base", group="base"), data=yearly[1:13,]) +
  expand_limits(y=0) + scale_x_continuous(minor_breaks=2011:2023)
print(figs$figure.20)
```


```{r connect-ai, eval=! exists("overlap.ai")}

t1 <- left_join(cpsc.ai %>% mutate(c.id=1:nrow(cpsc.ai), Year=lubridate::year(Report.Date), 
                                   City=tolower(City)) %>% filter(Lithium & Fire), 
                details %>% mutate(n.id=1:nrow(details), city=tolower(city), state_id=toupper(state_id), 
                                   year=lubridate::year(inc_date)), 
                by=join_by(City == city, st_abb == state_id, Year == year)) %>% 
      select(c.id, n.id, Report.Date, inc_date, City, st_abb, ZIP, zip5, fd_name, 
             Location, property, Severity, casualties, tod, mob_invol, room, heat, 
             ifi, fire_sprd, Description) %>% 
      filter(! is.na(n.id))

t2 <- left_join(cpsc.ai %>% mutate(c.id=1:nrow(cpsc.ai), Year=lubridate::year(Report.Date),
                                   City=tolower(City)) %>% filter(Lithium & Fire), 
                details %>% mutate(n.id=1:nrow(details), hdate=inc_date + 90, ldate=inc_date - 7,
                                   city=tolower(city), state_id=toupper(state_id)), 
                by=join_by(City == city, st_abb == state_id, Report.Date >= ldate, Report.Date <= hdate)) %>% 
      select(c.id, n.id, Report.Date, inc_date, City, st_abb, ZIP, zip5, fd_name, 
             Location, property, Severity, casualties, tod, mob_invol, room, heat, 
             ifi, fire_sprd, Description) %>% 
      filter(! is.na(n.id))

# The following 'valid' values are determined by going through all the potential
# matches by hand to see if they are valid matches. 
# 
t1.valid <- c(F, F, F, F, F, F, T,
              F, F, F, F, F, F, F,
              F, F, F, F, F, F, T,
              F, F, T, T, F, F, F,
              F, F, F, F, F, F, F,
              T, F, F, F, F, F, F,
              F, F, F, F, F, F, F)

t2.valid <- c(T, F, F, F, 
              F, T, F, T, 
              T, F, F, T, 
              F, F, F, F)

overlap.ai <- t2 %>% filter(t2.valid) %>% 
                     select(c.id, n.id, City, st_abb, ZIP, Report.Date, 
                            inc_date, Location, Severity)

overlap <- cbind(cpsc.ai %>% {.[(1:nrow(cpsc.ai))[! is.na(tmp)],
                                c("City", "st_abb", "ZIP", "Report.Date", 
                                  "Location", "Severity")]},
                 details[tmp[! is.na(tmp)],c(12:14,5,7)])
```

The AI-screened CPSC dataset was matched to the NFIRS dataset based on city,
state and date. If city and state matched and the dates were within specified
bounds, the records were considered candidate matches. The bounds used were
that the CPSC complaint was made not more than 90 days after the NFIRS report
and not more than seven days before. That produced 
`r cpsc.ai %>% filter(Lithium, Fire) %>% nrow()` records which were then
reviewed by hand to determine whether they were the same incident. A total of
`r overlap.ai %>% nrow()` overlapping records were identified.

```{r yearly}
yearly <- data.frame(Year=c(2011:2023, 0, -(1:5)), 
                     cpsc=c(cpsc.ai %>% mutate(year=lubridate::year(Report.Date)) %>% 
                              filter(year > 2010, year < 2024) %>% with(table(year)), rep(NA, 6)),
                     nfirs=c(details %>% with(table(year)), rep(NA, 6)),
                     overlap=c(overlap %>% mutate(year=lubridate::year(Report.Date)) %>%
                                 filter(year < 2024) %>% with(table(year)), rep(NA, 6)))

# `yearly` is structured so that the first 13 rows are the years from 2011-2023
# Year 0, then is the total of all years
# Year -1 to -4 are different versions of total number of injuries.
# Year -5 is the AI estimate.

yearly$cpsc[14]    <- sum(yearly$cpsc[1:13])
yearly$nfirs[14]   <- sum(yearly$nfirs[1:13])
yearly$overlap[14] <- sum(yearly$overlap[1:13])
yearly$cpsc[15:18] <- cpsc.wk %>% filter(Severity %in% c("Injury", "Death")) %>% nrow()
yearly$nfirs[15:18]<- details %>% filter(casualties == "Yes") %>% nrow()
yearly$overlap[15] <- overlap %>% filter(Severity %in% c("Injury", "Death")) %>% nrow()
yearly$overlap[16] <- overlap %>% filter(casualties == "Yes") %>% nrow()
yearly$overlap[17] <- overlap %>% filter(Severity %in% c("Injury", "Death"), 
                                         casualties == "Yes") %>% nrow()
yearly$overlap[18] <- overlap %>% filter(Severity %in% c("Injury", "Death") | 
                                           casualties == "Yes") %>% nrow()

yearly$cpsc[19] <- cpsc.ai %>% filter(Lithium, Fire) %>% nrow()
yearly$nfirs[19] <- sum(yearly$nfirs[1:13])
yearly$overlap[19] <- overlap.ai %>% nrow()

yearly$N  <- with(yearly, ((cpsc + 1) * (nfirs + 1) / (overlap + 1)) - 1)
yearly$se <- with(yearly, sqrt((N - cpsc) * (N - nfirs) / overlap))
yearly$lcl <- yearly$N + qnorm(0.025) * yearly$se
yearly$ucl <- yearly$N + qnorm(0.975) * yearly$se
```

The idea here is that if you have two (or more) independent samples of a 
population, you can exploit that to estimate the size of the population
provided you have some cases that can be identified in both samples.
The approach was pioneered by Petersen (C. Petersen, "The yearly immigration of young
plaice, into the Limfjord from the German Sea" *Rept. Danish Biol. Sta.* 1896, 6, 1-48). 
Here the model is estimated using Bayesian Markov Chain Monte Carlo techniques.

Let *N* be the size of the population, and $p_i$ be the probability of any 
individual from the population occurring in sample *i* 
(where $i \in \left\{\mathrm{cpsc}, \mathrm{nfirs}\right\}$). 
Then the size of sample *i* is distributed as:

$$S_i \sim Binom\left(N, p_i \right)$$

and the number of individual cases in ***both*** samples is distributed as:

$$S_{ij} \sim Binom\left(N, p_{cpsc} \cdot p_{nfirs} \right)$$

Model was estimated using STAN (Stan Development Team. 2025. Stan User’s Guide, 2.36. https://mc-stan.org).

Overlap in the cases were identified by matching city, state and year. While 
there may be some spurious overlaps given the approximate measures used to 
identify overlaps, cases are rare enough that overlaps are unlikely.

```{stan connect-stan, output.var="stan.con.model", eval=! exists("stan.con.model")}
data {
  int<lower=2>         N;        // Number of independent surveys
  matrix[N, N]         X;        // Counts. Only the lower-triangular part is used
}

transformed data {
  real     min_M;
  real     scale_M = 0;
  for(i in 2:N){
    for(j in 1:(i - 1)){
    scale_M += X[i,i] * X[j,j] / X[i,j];
  }}
  scale_M = 2 * scale_M / (N * (N - 1));
  
  min_M = max(diagonal(X)) / scale_M;
}

parameters {
  vector<lower=0, upper=1>[N]   p;
  real<lower=min_M>             M_hat;
}

transformed parameters {
  real    M = M_hat * scale_M;
}

model {
  real   q;
  real   val;
  for(i in 1:N){
    for(j in 1:i){
      if(i == j){
        q = p[i];
      } else {
        q = p[i] * p[j];
      }
      val = X[i,j];
      
      target += lgamma(M + 1) - lgamma(val + 1) - lgamma(M - val + 1)
                + val * log(q) + (M - val) * log(1 - q);
  }}
}

generated quantities {
}
```


```{stan connect-stan-r1, output.var="stan.con.mdl1", eval=! exists("stan.con.mdl1")}
// This differs from the version above in that it computes a slope over time
// instead of computing estimated values for each year separately.
data {
  int<lower=1>          T;       // Number of time periods
  vector[T]             S1;      // Sample 1 Counts
  vector[T]             S2;      // Sample 2 Counts
  vector[T]             So;      // Sample Overlap Counts
  int<lower=0, upper=1> slope1;  // 0: No; 1: Yes
  int<lower=0, upper=1> slope2;  // 0: No; 1: Yes
}

transformed data {

  matrix[T,3]    dats = append_col(append_col(S1, S2), So);
  
  real     min_N;
  real     scale_N = 0;
  for(i in 1:T){
    scale_N += S1[i] * S2[i] / So[i];
  }
  scale_N = scale_N / T;
  
  min_N = max(append_row(S1, S2)) / scale_N;
}

parameters {
  vector[1 + slope1]       p1_lgt;
  vector[1 + slope2]       p2_lgt;
  real                     r;
  real<lower=min_N>        N_hat;
}

transformed parameters {
  vector[T]    N;
  vector[T]    p1;
  vector[T]    p2;
  for(i in 1:T){
    N[i] = N_hat * scale_N * exp((i - 1) * r);
    if(slope1 == 1){
      p1[i] = inv_logit(p1_lgt[1] + (i - 1) * p1_lgt[2]);
    } else {
      p1[i] = inv_logit(p1_lgt[1]);
    }
    if(slope2 == 1){
      p2[i] = inv_logit(p2_lgt[1] + (i - 1) * p2_lgt[2]);
    } else {
      p2[i] = inv_logit(p2_lgt[1]);
    }
  }
}

model {
  matrix[T, 3]   p_hat = append_col(append_col(p1, p2), p1 .* p2);
  for(i in 1:T){
    for(j in 1:3){
      target += lgamma(N[i] + 1) - lgamma(dats[i,j] + 1) - lgamma(N[i] - dats[i,j] + 1)
              + dats[i,j] * log(p_hat[i,j]) + (N[i] - dats[i,j]) * log(1 - p_hat[i,j]);
  }}
}

generated quantities {
}
```

```{r con-stan-run, eval=! exists("stan.con.fit")}
stan.con.fit <- list()
st.data <- list(N=2, X=NA)
for(i in 1:nrow(yearly)){
  st.data$X <- with(yearly, matrix(c(cpsc[i], overlap[i], 0, nfirs[i]), nrow=2))
  stan.con.fit[[i]] <- sampling(stan.con.model, data=st.data, refresh=0, iter=10000, warmup=1000)
}
```

```{r con-stan-run1, eval=! exists("stan.con.fit1")}
st.data1 <- yearly %>% filter(Year > 1000) %>% 
            with(list(T=length(Year),
                      S1=cpsc,
                      S2=nfirs,
                      So=overlap,
                      slope1=0,
                      slope2=1))

inits <- list(list(p1_lgt=c(-3, 0), p2_lgt=c(-3, 0), r=0.1, M_hat=0.5),
              list(p1_lgt=c(-3, 0), p2_lgt=c(-3, 0), r=0.1, M_hat=0.5),
              list(p1_lgt=c(-3, 0), p2_lgt=c(-3, 0), r=0.1, M_hat=0.5),
              list(p1_lgt=c(-3, 0), p2_lgt=c(-3, 0), r=0.1, M_hat=0.5))

inits <- function(){
  list(p1_lgt=as.array(rnorm(1, -3, .25)), 
       p2_lgt=rnorm(2, c(-3, 0), 0.25),
       r=rnorm(1, 0, 0.2),
       M_hat=rnorm(1, 1, 0.25))
}

stan.con.fit1 <- sampling(stan.con.mdl1, data=st.data1, init=inits, 
                          iter=2000, warmup=1000)
```

```{r con-stan-res}
yearly.stan <- yearly[,1:4]

ww <- NULL
vv <- NULL
for(i in 1:length(stan.con.fit)){
  ww <- rbind(ww, rstan::extract(stan.con.fit[[i]], "M")[[1]] %>% 
        { c(mean(.), sd(.), quantile(., c(0.025, 0.5, 0.975))) })
  vv <- cbind(vv, rstan::extract(stan.con.fit[[i]], "p")[[1]] %>% 
        apply(2, function(x) c(mean(x), sd(x), quantile(x, c(0.025, 0.5, 0.975)))))
}

ww <- ww[,-4] %>% as.data.frame() %>% 
      setNames(c("N", paste0("N.", c("se", "lcl", "ucl"))))
v1 <- vv[-4,1:ncol(vv) %% 2 == 1] %>% t() %>% as.data.frame() %>% 
      setNames(c("p.cpsc", paste0("p.cpsc.", c("se", "lcl", "ucl"))))
v2 <- vv[-4,1:ncol(vv) %% 2 == 0] %>% t() %>% as.data.frame() %>% 
      setNames(c("p.nfirs", paste0("p.nfirs.", c("se", "lcl", "ucl"))))

yearly.stan <- cbind(yearly.stan, ww, v1, v2)
```


```{r con-stan-res1}
yearly.stan1 <- yearly[1:13,1:4]

yearly.stan1 <- cbind(yearly.stan1, 
                      stan.con.fit1 %>% { rstan::extract(., "M")[[1]] } %>% 
                        apply(2, function(x) c(mean(x), sd(x), 
                                               quantile(x, c(0.5, 0.025, 0.975)))) %>%
                        t() %>% as.data.frame() %>% 
                        setNames(c("N", "N.se", "N.median", "N.lcl", "N.ucl")),
                      stan.con.fit1 %>% { rstan::extract(., "p1")[[1]] } %>% 
                        apply(2, function(x) c(mean(x), sd(x), 
                                               quantile(x, c(0.5, 0.025, 0.975)))) %>%
                        t() %>% as.data.frame() %>% 
                        setNames(c("p.cpsc", "p.cpsc.se", "p.cpsc.median", 
                                   "p.cpsc.lcl", "p.cpsc.ucl")),
                      stan.con.fit1 %>% { rstan::extract(., "p2")[[1]] } %>% 
                        apply(2, function(x) c(mean(x), sd(x), 
                                               quantile(x, c(0.5, 0.025, 0.975)))) %>%
                        t() %>% as.data.frame() %>% 
                        setNames(c("p.nfirs", "p.nfirs.se", "p.nfirs.median",
                                   "p.nfirs.lcl", "p.nfirs.ucl")))

slopey <- rbind(stan.con.fit1 %>% {rstan::extract(., "r")[[1]] } %>% {exp(.) - 1} %>% 
                    {c(mean(.), sd(.), quantile(., c(0.5, 0.025, 0.975))) },
                stan.con.fit1 %>% {rstan::extract(., "p1_lgt")[[1]][,2] } %>% 
                    {c(mean(.), sd(.), quantile(., c(0.5, 0.025, 0.975))) },
                stan.con.fit1 %>% {rstan::extract(., "p2_lgt")[[1]][,2] } %>% 
                    {c(mean(.), sd(.), quantile(., c(0.5, 0.025, 0.975))) }) %>%
          as.data.frame() %>% setNames(c("mean", "se", "median", "lcl", "ucl"))

slopey <- cbind(var=c("r", "p_cpsc", "p_nfirs"), slopey)
```

There were `r yearly.stan[14,2]` relevant cases in the CPSC data and 
`r yearly.stan[14, 3]` relevant cases from the NFIRS data, with an overlap 
of `r yearly.stan[14,4]` cases. This implies a total population of battery
fires of approximately `r yearly.stan[14,5] %>% round(-3) %>% formatC(format="d")` 
(confidence interval: 
`r yearly.stan[14,7] %>% round(-3) %>% formatC(format="d")` - 
`r yearly.stan[14,8] %>% round(-3) %>% formatC(format="d")`).

Using the AI-screened data set, there were `r yearly.stan[19,2]` relevant cases 
in the CPSC data, `r yearly.stan[19, 3]` relevant cases from the NFIRS data, 
and an overlap of `r yearly.stan[19,4]` cases. This implies a total population of battery
fires of approximately `r yearly.stan[19,5] %>% round(-3) %>% formatC(format="d")` 
(confidence interval: 
`r yearly.stan[19,7] %>% round(-3) %>% formatC(format="d")` - 
`r yearly.stan[19,8] %>% round(-3) %>% formatC(format="d")`).

The chart below shows the estimated number of battery fires by year between 
2011 and 2023.

```{r overlap-graph}
figs$figure.21 <- yearly.stan %>% filter(Year > 2000) %>% ggplot() + 
          aes(x=Year, y=N, ymin=N.lcl, ymax=N.ucl) + 
          geom_point() + geom_line() + geom_ribbon(alpha=0.3) +
          geom_line(data=yearly.stan1, color="blue") + 
          geom_ribbon(data=yearly.stan1, color="blue", alpha=0.3) + 
          scale_x_continuous(breaks=2010 + 5 * 0:3, minor_breaks = 2010:2025) +
          expand_limits(x=2010) + theme_bw()
print(figs$figure.21)
```


```{r p-nfirs-graph}
figs$figure.22 <- yearly.stan %>% filter(Year > 2000) %>% ggplot() + 
          aes(x=Year, y=p.nfirs, ymin=p.nfirs.lcl, ymax=p.nfirs.ucl) + 
          geom_point() + geom_line() + geom_ribbon(alpha=0.3) +
          geom_line(data=yearly.stan1, color="blue") + 
          geom_ribbon(data=yearly.stan1, color="blue", alpha=0.3) +
          scale_x_continuous(breaks=2010 + 5 * 0:3, minor_breaks = 2010:2025)
print(figs$figure.22)
```


```{r p-cpsc-graph}
figs$figure.23 <- yearly.stan %>% filter(Year > 2000) %>% ggplot() + 
          aes(x=Year, y=p.cpsc, ymin=p.cpsc.lcl, ymax=p.cpsc.ucl) + 
          geom_point() + geom_line() + geom_ribbon(alpha=0.3) +
          geom_line(data=yearly.stan1, color="blue") + 
          geom_ribbon(data=yearly.stan1, color="blue", alpha=0.3) +
          scale_x_continuous(breaks=2010 + 5 * 0:3, minor_breaks = 2010:2025)
print(figs$figure.23)
```

Growth rate in the estimated number of battery fires is 
`r (slopey$mean[1] * 100) %>% formatC(digits=2)`&nbsp;% 
(95&nbsp;% confidence interval: `r (slopey$lcl[1] * 100) %>% formatC(digits=2)` - 
`r (slopey$ucl[1] * 100) %>% formatC(digits=2)`&nbsp;%) per year. The sampling 
rate for battery fires in the CPSC data does not change significantly over time, 
while the sampling rate for battery fires in the NFIRS data increases from 
`r (yearly.stan1$p.nfirs[1] * 100) %>% formatC(digits=2)`&nbsp;% 
(95&nbsp;% confidence interval: 
`r (yearly.stan1$p.nfirs.lcl[1] * 100) %>% formatC(digits=2)` - 
`r (yearly.stan1$p.nfirs.ucl[1] * 100) %>% formatC(digits=2)`&nbsp;%) in 2011 to 
`r (yearly.stan1$p.nfirs[13] * 100) %>% formatC(digits=2)`&nbsp;%  
(95&nbsp;% confidence interval:
`r (yearly.stan1$p.nfirs.lcl[13] * 100) %>% formatC(digits=2)` - 
`r (yearly.stan1$p.nfirs.ucl[13] * 100) %>% formatC(digits=2)`&nbsp;%) in 2023.

## Injuries

There is not enough data to estimate yearly values for injuries. Overlaps for 
injuries are more difficult since injuries are not always reported. An example 
of that is a fire in Washington that is in both data sets, where a minor injury 
is reported in the CPSC data set but is not reported in the NFIRS data set.
Number of injuries depends on how overlapping number of injuries is counted. 
If an injury is counted if it is recorded in *both* data sets, then there were
`r yearly.stan$N[17] %>% round(-2) %>% formatC(format="d", big.mark = "&nbsp;")`
(95&nbsp;% confidence interval: 
`r yearly.stan$N.lcl[17] %>% round(-2) %>% formatC(format="d")` - 
`r yearly.stan$N.ucl[17] %>% round(-2) %>% formatC(format="d", big.mark = "&nbsp;")`) 
injuries related to battery fires between 2011 and 2023. If an injury is counted if 
it is recorded in *either* data set, then there were 
`r yearly.stan$N[18] %>% round(-2) %>% formatC(format="d", big.mark = "&nbsp;")`
(95&nbsp;% confidence interval: 
`r yearly.stan$N.lcl[18] %>% round(-2) %>% formatC(format="d")` - 
`r yearly.stan$N.ucl[18] %>% round(-2) %>% formatC(format="d", big.mark = "&nbsp;")`) 
injuries related to battery fires between 2011 and 2023. 

```{r cleanup-04}
# Save the relevant file, as below. Then delete everything if you are working in bookdown.
# That means that a `bookdown` flag needs to be set somehow. And then unset at the end.
save(list=intersect(keep2, ls()), file=paste0(intmdt_dir, "recapture.RData"))
rm(list=setdiff(ls(), keep))
```