---
title: "Batteries"
subtitle: "Capture-Recapture Estimates"
output: html_document
date: "2024-12-13"
---

# Combined NFIRS and CPSC

```{r setup-04, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)
#
# For now, keep the old methods in, but soon, delete them and leave
# just the ai-screened set.
#
library(magrittr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(rstan)

keep <- c(ls(), "figs")

if(! exists("csv_dir")){
  csv_dir <- "csv/"
}
if(! exists("intmdt_dir")){
  intmdt_dir <- "rdata/"
}

source("load.files.R")
if(
  load.data(final=c("recapture.RData", "stan.models.RData"), 
            data =c("recapture.data.RData", "cpsc.inc.RData", "NFIRS.RData", "stan.models.RData"), 
            dir=intmdt_dir) == 1
  ) stop("The data files needed for the Recapture analysis are not present!")

rm(load.data)

keep2 <- list(recapture=c("details", "cpsc.ai", "overlap.ai", 
                          "yearly", "stan.con.fit", "stan.con.fit1", "yearly.stan")
              )
if(! exists("figs")) figs <- list()
```

## Connect to NFIRS

```{r connect-ai, eval=! exists("overlap.ai")}

t1 <- left_join(cpsc.ai %>% mutate(c.id=1:nrow(cpsc.ai), Year=lubridate::year(Report.Date), 
                                   City=tolower(City)) %>% filter(Lithium & Fire), 
                details %>% mutate(n.id=1:nrow(details), city=tolower(city), state_id=toupper(state_id), 
                                   year=lubridate::year(inc_date)), 
                by=join_by(City == city, st_abb == state_id, Year == year)) %>% 
      select(c.id, n.id, Report.Date, inc_date, City, st_abb, ZIP, zip5, fd_name, 
             Location, property, Severity, casualties, tod, mob_invol, room, heat, 
             ifi, fire_sprd, Description) %>% 
      filter(! is.na(n.id))

t2 <- left_join(cpsc.ai %>% mutate(c.id=1:nrow(cpsc.ai), Year=lubridate::year(Report.Date),
                                   City=tolower(City)) %>% filter(Lithium & Fire), 
                details %>% mutate(n.id=1:nrow(details), hdate=inc_date + 90, ldate=inc_date - 7,
                                   city=tolower(city), state_id=toupper(state_id)), 
                by=join_by(City == city, st_abb == state_id, Report.Date >= ldate, Report.Date <= hdate)) %>% 
      select(c.id, n.id, Report.Date, inc_date, City, st_abb, ZIP, zip5, fd_name, 
             Location, property, Severity, casualties, tod, mob_invol, room, heat, 
             ifi, fire_sprd, Description) %>% 
      filter(! is.na(n.id))

# The following 'valid' values are determined by going through all the potential
# matches by hand to see if they are valid matches. 
# 
t1.valid <- c(F, F, F, F, F, F, T,
              F, F, F, F, F, F, F,
              F, F, F, F, F, F, T,
              F, F, T, T, F, F, F,
              F, F, F, F, F, F, F,
              T, F, F, F, F, F, F,
              F, F, F, F, F, F, F)

t2.valid <- c(T, F, F, F, 
              F, T, F, T, 
              T, F, F, T, 
              F, F, F, F)

overlap.ai <- t2 %>% filter(t2.valid) %>% 
                     select(c.id, n.id, City, st_abb, ZIP, Report.Date, 
                            inc_date, Location, Severity)
```

The AI-screened CPSC dataset was matched to the NFIRS dataset based on city,
state and date. If city and state matched and the dates were within specified
bounds, the records were considered candidate matches. The bounds used were
that the CPSC complaint was made not more than 90 days after the NFIRS report
and not more than seven days before. That produced 
`r cpsc.ai %>% filter(Lithium, Fire) %>% nrow()` records which were then
reviewed by hand to determine whether they were the same incident. A total of
`r overlap.ai %>% nrow()` overlapping records were identified.

```{r yearly}
yearly <- data.frame(Year=c(2011:2023, 0), 
                     cpsc=c(cpsc.ai %>% mutate(year=lubridate::year(Report.Date)) %>% 
                              filter(year > 2010, year < 2024, Lithium, Fire) %>% with(table(year)),
                            NA),
                     nfirs=c(details %>% with(table(year)), NA)) %>%
          left_join(overlap.ai %>% mutate(Year=lubridate::year(Report.Date)) %>%
                                group_by(Year) %>% summarize(overlap=n()))

# `yearly` is structured so that the first 13 rows are the years from 2011-2023
# Year 0, then is the total of all years

yearly$cpsc[14]    <- sum(yearly$cpsc[1:13], na.rm=T)
yearly$nfirs[14]   <- sum(yearly$nfirs[1:13], na.rm=T)
yearly$overlap[14] <- sum(yearly$overlap[1:13], na.rm=T)

yearly$N  <- with(yearly, ((cpsc + 1) * (nfirs + 1) / (overlap + 1)) - 1)
yearly$se <- with(yearly, sqrt((N - cpsc) * (N - nfirs) / overlap))
yearly$lcl <- yearly$N + qnorm(0.025) * yearly$se
yearly$ucl <- yearly$N + qnorm(0.975) * yearly$se
```

The idea here is that if you have two (or more) independent samples of a 
population, you can exploit that to estimate the size of the population
provided you have some cases that can be identified in both samples.
The approach was pioneered by Petersen (C. Petersen, "The yearly immigration of young
plaice, into the Limfjord from the German Sea" *Rept. Danish Biol. Sta.* 1896, 6, 1-48). 
Here the model is estimated using Bayesian Markov Chain Monte Carlo techniques.

Let *N* be the size of the population, and $p_i$ be the probability of any 
individual from the population occurring in sample *i* 
(where $i \in \left\{\mathrm{cpsc}, \mathrm{nfirs}\right\}$). 
Then the size of sample *i* is distributed as:

$$S_i \sim Binom\left(N, p_i \right)$$

and the number of individual cases in ***both*** samples is distributed as:

$$S_{ij} \sim Binom\left(N, p_{cpsc} \cdot p_{nfirs} \right)$$

Model was estimated using STAN (Stan Development Team. 2025. Stan User’s Guide, 
2.36. https://mc-stan.org).

Overlap in the cases were identified by matching city, state and year. While 
there may be some spurious overlaps given the approximate measures used to 
identify overlaps, cases are rare enough that overlaps are unlikely.

```{stan connect-stan, output.var="stan.con.model", eval=! exists("stan.con.model")}
data {
  int<lower=2>         N;        // Number of independent surveys
  matrix[N, N]         X;        // Counts. Only the lower-triangular part is used
}

transformed data {
  real     min_M;
  real     scale_M = 0;
  for(i in 2:N){
    for(j in 1:(i - 1)){
    scale_M += X[i,i] * X[j,j] / X[i,j];
  }}
  scale_M = 2 * scale_M / (N * (N - 1));
  
  min_M = max(diagonal(X)) / scale_M;
}

parameters {
  vector<lower=0, upper=1>[N]   p;
  real<lower=min_M>             M_hat;
}

transformed parameters {
  real    M = M_hat * scale_M;
}

model {
  real   q;
  real   val;
  for(i in 1:N){
    for(j in 1:i){
      if(i == j){
        q = p[i];
      } else {
        q = p[i] * p[j];
      }
      val = X[i,j];
      
      target += lgamma(M + 1) - lgamma(val + 1) - lgamma(M - val + 1)
                + val * log(q) + (M - val) * log(1 - q);
  }}
}

generated quantities {
}
```


```{stan connect-stan-r1, output.var="stan.con.mdl1", eval=! exists("stan.con.mdl1")}
// This differs from the version above in that it computes a slope over time
// instead of computing estimated values for each year separately.
data {
  int<lower=1>          T;       // Number of time periods
  vector[T]             S1;      // Sample 1 Counts
  vector[T]             S2;      // Sample 2 Counts
  vector[T]             So;      // Sample Overlap Counts
  int<lower=0, upper=1> slope1;  // 0: No; 1: Yes
  int<lower=0, upper=1> slope2;  // 0: No; 1: Yes
}

transformed data {

  matrix[T,3]    dats = append_col(append_col(S1, S2), So);
  
  real     min_N;
  real     scale_N = 0;
  for(i in 1:T){
    scale_N += S1[i] * S2[i] / So[i];
  }
  scale_N = scale_N / T;
  
  min_N = max(append_row(S1, S2)) / scale_N;
}

parameters {
  vector[1 + slope1]       p1_lgt;
  vector[1 + slope2]       p2_lgt;
  real                     r;
  real<lower=min_N>        N_hat;
}

transformed parameters {
  vector[T]    N;
  vector[T]    p1;
  vector[T]    p2;
  for(i in 1:T){
    N[i] = N_hat * scale_N * exp((i - 1) * r);
    if(slope1 == 1){
      p1[i] = inv_logit(p1_lgt[1] + (i - 1) * p1_lgt[2]);
    } else {
      p1[i] = inv_logit(p1_lgt[1]);
    }
    if(slope2 == 1){
      p2[i] = inv_logit(p2_lgt[1] + (i - 1) * p2_lgt[2]);
    } else {
      p2[i] = inv_logit(p2_lgt[1]);
    }
  }
}

model {
  matrix[T, 3]   p_hat = append_col(append_col(p1, p2), p1 .* p2);
  for(i in 1:T){
    for(j in 1:3){
      target += lgamma(N[i] + 1) - lgamma(dats[i,j] + 1) - lgamma(N[i] - dats[i,j] + 1)
              + dats[i,j] * log(p_hat[i,j]) + (N[i] - dats[i,j]) * log(1 - p_hat[i,j]);
  }}
}

generated quantities {
}
```

```{r con-stan-run, eval=! exists("stan.con.fit")}
stan.con.fit <- list()

st.data <- list(S1=NA,
                S2=NA,
                S12=NA,
                alpha=0.1,
                scale_p=1)

for(i in 1:nrow(yearly)){
  if(! is.na(yearly$overlap[i])){
    nm <- as.character(yearly$Year[i])
    st.data$S1  <- yearly$cpsc[i]
    st.data$S2  <- yearly$nfirs[i]
    st.data$S12 <- yearly$overlap[i]
    stan.con.fit[[nm]] <- sampling(stan.model, data=st.data, refresh=0, iter=10000, warmup=1000)
  }
}
```

```{r con-stan-run1, eval=! exists("stan.con.fit1")}
ndx <- (1:nrow(yearly))[yearly$Year > 2016 & yearly$Year < 2024]
st.data1 <- yearly[ndx,] %>% 
            with(list(T=length(Year),
                      S1=cpsc,
                      S2=nfirs,
                      So=ifelse(is.na(overlap), 0, overlap),
                      slope1=0,
                      slope2=0))

stan.con.fit1 <- sampling(stan.mode1, data=st.data1, iter=2000, warmup=1000)
```

```{r con-stan-res}
yearly.stan <- yearly[,1:4]
yearly.stan$N     <- NA
yearly.stan$N.se  <- NA
yearly.stan$N.lcl <- NA
yearly.stan$N.med <- NA
yearly.stan$N.ucl <- NA

for(i in names(stan.con.fit)){
  yr <- match(as.integer(i), yearly.stan$Year)
  yearly.stan[yr, 5:9] <- rstan::extract(stan.con.fit[[i]], "N")[[1]] %>% 
                          { c(mean(.), sd(.), quantile(., c(0.025, 0.5, 0.975))) }
}

```


```{r con-stan-res1}
ndx <- (1:nrow(yearly))[yearly$Year > 2016 & yearly$Year < 2024]
yearly.stan1 <- yearly[ndx,1:4]

yearly.stan1 <- cbind(yearly.stan1, 
                      stan.con.fit1 %>% { rstan::extract(., "N")[[1]] } %>% 
                        apply(2, function(x) c(mean(x), sd(x), 
                                               quantile(x, c(0.5, 0.025, 0.975)))) %>%
                        t() %>% as.data.frame() %>% 
                        setNames(c("N", "N.se", "N.median", "N.lcl", "N.ucl")),
                      stan.con.fit1 %>% { rstan::extract(., "p1")[[1]] } %>% 
                        apply(2, function(x) c(mean(x), sd(x), 
                                               quantile(x, c(0.5, 0.025, 0.975)))) %>%
                        t() %>% as.data.frame() %>% 
                        setNames(c("p.cpsc", "p.cpsc.se", "p.cpsc.median", 
                                   "p.cpsc.lcl", "p.cpsc.ucl")),
                      stan.con.fit1 %>% { rstan::extract(., "p2")[[1]] } %>% 
                        apply(2, function(x) c(mean(x), sd(x), 
                                               quantile(x, c(0.5, 0.025, 0.975)))) %>%
                        t() %>% as.data.frame() %>% 
                        setNames(c("p.nfirs", "p.nfirs.se", "p.nfirs.median",
                                   "p.nfirs.lcl", "p.nfirs.ucl")))

slopey <- stan.con.fit1 %>% {rstan::extract(., "r")[[1]] } %>% {exp(.) - 1} %>% 
                    {c(mean=mean(.), se=sd(.), quantile(., c(0.5, 0.025, 0.975))) }
```

There were `r yearly.stan[14,2]` relevant cases in the CPSC data and 
`r yearly.stan[14, 3]` relevant cases from the NFIRS data, with an overlap 
of `r yearly.stan[14,4]` cases. This implies a total population of battery
fires of approximately `r yearly.stan[14,5] %>% round(-3) %>% formatC(format="d")` 
(confidence interval: 
`r yearly.stan[14,7] %>% round(-3) %>% formatC(format="d")` - 
`r yearly.stan[14,8] %>% round(-3) %>% formatC(format="d")`).

The chart below shows the estimated number of battery fires by year between 
2011 and 2023.

```{r overlap-graph}
figs$figure.21 <- yearly.stan %>% filter(Year > 2000) %>% ggplot() + 
          aes(x=Year, y=N, ymin=N.lcl, ymax=N.ucl) + 
          geom_point() + geom_errorbar() +
          geom_line(data=yearly.stan1, color="blue") + 
          geom_ribbon(data=yearly.stan1, color="blue", alpha=0.3) + 
          scale_x_continuous(breaks=2010 + 5 * 0:3, minor_breaks = 2010:2025) +
          expand_limits(x=2010) + theme_bw()
print(figs$figure.21)
```


```{r p-nfirs-graph, eval=F}
figs$figure.22 <- yearly.stan %>% filter(Year > 2000) %>% ggplot() + 
          aes(x=Year, y=p.nfirs, ymin=p.nfirs.lcl, ymax=p.nfirs.ucl) + 
          geom_point() + geom_line() + geom_ribbon(alpha=0.3) +
          geom_line(data=yearly.stan1, color="blue") + 
          geom_ribbon(data=yearly.stan1, color="blue", alpha=0.3) +
          scale_x_continuous(breaks=2010 + 5 * 0:3, minor_breaks = 2010:2025)
print(figs$figure.22)
```


```{r p-cpsc-graph, eval=F}
figs$figure.23 <- yearly.stan %>% filter(Year > 2000) %>% ggplot() + 
          aes(x=Year, y=p.cpsc, ymin=p.cpsc.lcl, ymax=p.cpsc.ucl) + 
          geom_point() + geom_line() + geom_ribbon(alpha=0.3) +
          geom_line(data=yearly.stan1, color="blue") + 
          geom_ribbon(data=yearly.stan1, color="blue", alpha=0.3) +
          scale_x_continuous(breaks=2010 + 5 * 0:3, minor_breaks = 2010:2025)
print(figs$figure.23)
```

Growth rate in the estimated number of battery fires is 
`r (slopey["mean"] * 100) %>% formatC(digits=2)`&nbsp;% 
(95&nbsp;% confidence interval: `r (slopey["lcl"] * 100) %>% formatC(digits=2)` - 
`r (slopey["ucl"] * 100) %>% formatC(digits=2)`&nbsp;%) per year.

```{r cleanup-04}
# Save the relevant file, as below. Then delete everything if you are working in bookdown.
# That means that a `bookdown` flag needs to be set somehow. And then unset at the end.
for(i in names(keep2)){
  save(list=intersect(keep2[[i]], ls()), file=paste0(intmdt_dir, i, ".RData"))
}
rm(list=setdiff(ls(), keep))
```