---
title: "Batteries"
subtitle: "Battery Energy Storage Systems"
output: html_document
date: "2024-12-13"
---

# BESS

```{r setup-06, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE)

library(magrittr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(rstan)

keep <- ls()

if(! exists("csv_dir")){
  csv_dir <- "csv/"
}
if(! exists("intmdt_dir")){
  intmdt_dir <- "rdata/"
}

ff <- list.files(intmdt_dir) %>% grep("RData", ., ignore.case=T, value=T)
if("bess.rdata" %in% tolower(ff)){
  fg <- "bess.rdata" %>% match(tolower(ff)) %>% { ff[.] }
  load(paste0(intmdt_dir, fg))
}
rm(ff, fg)

 keep2 <- c("bess", "bess.model", "figs")
```

```{r bess}
# This table was downloaded from:
# https://storagewiki.epri.com/resources/assets/BESS_Failure_Database/Failure_DB_List.csv

bess <- paste0(csv_dir, "BESS Failure.csv") %>% read.csv()
```

```{r bess-plot}
# Data source for this is recorded above

figs$figure.40 <- merge(expand.grid(quarter=(1:4 - 0.5) / 4, year=2011:2025) %>% 
                          mutate(q=year + quarter),
                        bess %>% 
        mutate(q=lubridate::year(Event.Date) + (lubridate::quarter(Event.Date)-0.5)/4) %>% group_by(q) %>%
                 summarize(n=n()), all.x=T) %>% 
  ggplot() + aes(x=q, y=n) + geom_col()
print(figs$figure.40)
```


```{r bess-est}
bess.tab <- merge(expand.grid(quarter=(1:4 - 0.5) / 4, year=2011:2025) %>% 
                    mutate(q=year + quarter),
                  bess %>% mutate(q=lubridate::year(Event.Date) +
                                    (lubridate::quarter(Event.Date)-0.5)/4) %>% 
                    group_by(q) %>% summarize(n=n()), all.x=T) %>% 
  filter(year >= 2018, year < 2025) %>% mutate(n=ifelse(is.na(n), 0, n))

fn <- function(lambda, counts){
   sum(counts * dpois(0:7, lambda, log=T))
}

bess.tab1 <- bess.tab %>% with(table(n)) %>% as.vector() %>% data.frame(n=0:7, count=.)

bess.model <- list(trend=glm(n ~ I(q - 2025), data=bess.tab, family=quasipoisson),
                   no_trend=glm(n ~ 1, data=bess.tab, family=quasipoisson),
                   maxlik=maxLik(fn, start=c(lambda=3), counts=bess.tab1$count))

bess.tab1$model <- with(bess.tab1, sum(count) * dpois(n, bess.model$maxlik$estimate))

# bess.tab1 %>% knitr::kable()
```

This table (which is no longer there) and the next plot are probably superfluous. 
They just show that the
model fits well. The model is that events follow a Poisson process with 
`r formatC(bess.model$maxlik$estimate, digits=3)` &pm; 
`r formatC(qnorm(0.975) * as.vector(sqrt(-1/bess.model$maxlik$hessian)), digits=2)`. 
The value is the best fit, estimated using maximum likelihood.

I ran two more models, both glm models, one of which replicated the maximum
likelihood model, and the second of which added a trend. The constant glm model
gave the same results as the maximum likelihood model, and the trend was not
statistically significant.

Number of incidents over time is shown in the figure below.

```{r bess.time}
figs$figure.41 <- bess.tab %>% ggplot() + aes(x=q, y=n) + geom_point() + geom_line() +
             xlab("Time") + ylab("Count") + scale_x_continuous(breaks=2010:2025) + 
             theme_bw()
print(figs$figure.41)
```

```{r bess-plot2}
figs$figure.42 <- bess.tab1 %>% ggplot() + aes(x=n, y=count) + geom_col() + geom_line(aes(y=model))
print(figs$figure.42)
```

```{r cleanup-06}
save(list=intersect(keep2, ls()), file=paste0(intmdt_dir, "bess.RData"))
rm(list=setdiff(ls(), keep))
```